---
title: "Bounded scientific execution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bounded scientific execution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Whole-brain matrices can exceed memory even when the intended statistic is
simple. Bounded execution makes memory, materialization, and iteration limits
explicit at the scientific operation that consumes them. A resource budget
changes whether an operation may run; it must not silently change the
statistical method.

Resource limits stay attached to scientific operations rather than exposing
a second generic workflow engine.

## Streaming summaries

```{r}
backing <- cbind(signal = seq_len(12), predictor = 0:11)
x <- ngeo_point(
  cbind(x = 0:11, y = 0),
  values = backing,
  measures = rbind(
    ngeo_measure(support_behavior = "intensive"),
    ngeo_measure(support_behavior = "intensive")
  )
)
ngeo_stream_summary(x, chunk_size = 4)
ngeo_stream_lm(x, "signal", "predictor", chunk_size = 4)
```

## Bounded support change

Change of support consumes its sparse operator directly and checks the
declared materialization budget before computing values.

```{r}
map <- ngeo_atlas_map(
  x,
  rep(c("A", "B", "C"), each = 4),
  source_support = rep(1, 12)
)
target <- aggregate_to(
  x,
  map$target,
  map,
  budget = ngeo_resource_budget(
    memory_bytes = 1024^2,
    materialized_elements = 6
  )
)
target
```

## Scope

Auditable supported workflows use replay manifests; atomic filesystem
publication remains an internal responsibility of the format writers.
