---
title: "Bounded and resumable execution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bounded and resumable execution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

neurogeo 2.6 keeps the core representation unchanged: one spatial domain and
one strictly aligned values block. Bounded execution controls how that object
is processed.

```{r}
backing <- cbind(signal = seq_len(12), predictor = 0:11)
delayed <- ngeo_delayed_values(
  function(rows, columns) backing[rows, columns, drop = FALSE],
  dim(backing),
  map_names = colnames(backing),
  source = "documented callback"
)
x <- ngeo_points(
  cbind(x = 0:11, y = 0),
  values = delayed,
  measures = rbind(
    ngeo_measure(spatial_semantics = "intensive"),
    ngeo_measure(spatial_semantics = "intensive")
  )
)
ngeo_stream_summary(x, chunk_size = 4)
ngeo_stream_lm(x, "signal", "predictor", chunk_size = 4)
```

Block support operations consume sparse blocks directly.

```{r}
map <- ngeo_atlas_map(
  x,
  rep(c("A", "B", "C"), each = 4),
  source_support = rep(1, 12)
)
blocked <- ngeo_block_support_map(map, 1, 3)
target <- ngeo_change_support_block(
  x,
  map$target,
  blocked,
  budget = ngeo_resource_budget(blocks = 20)
)
ngeo_block_diagnostics(blocked)
```

Execution checkpoints are identity-bound. A changed domain, operator,
semantics, or parameter identity must use a new plan.

```{r}
checkpoint <- tempfile(fileext = ".json")
plan <- ngeo_execution_plan(
  "square",
  as.list(1:3),
  function(task, index) task^2,
  identity = list(domain_hash = ngeo_domain_hash(x)),
  checkpoint = checkpoint
)
ngeo_execute(plan)
```

Use `ngeo_atomic_write()` for final outputs. If the writer fails, the final
path is not created.
