---
title: "Bounded execution on vertex and DK case-control data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bounded execution on vertex and DK case-control data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
vertex <- ngeo_tutorial_vertex_case_control(n_per_group=5)
```

## The scientific map does not change when execution is chunked

The 672-by-10 vertex matrix contains five HC and five SCZ subjects. In a real
32k or 164k surface cohort, the same element-by-subject contract can exceed
memory. A resource budget changes whether an operation may run; it must never
silently change the estimator.

```{r brain, fig.width=9, fig.height=4.5, fig.alt="Vertex-level case-control cortical thickness contrast", fig.cap="Chunking changes execution, not the spatial map or its support."}
plot(vertex$difference, chart="flat", show_edges=FALSE, palette="Blue-Red 3")
```

## Stream summaries over cortical vertices

```{r streaming}
streamed <- ngeo_stream_summary(vertex$surface, chunk_size=96)
streamed
direct <- apply(ngeo_values(vertex$surface), 2L, function(value) {
  c(mean=mean(value), sd=stats::sd(value))
})
all.equal(unname(streamed$mean), unname(direct["mean", ]))
```

The chunk size is an execution parameter. Element order, subject layers,
measurement units, and the numerical summary are unchanged.

## Bound a sparse support change

```{r support}
map <- ngeo_atlas_map(
  vertex$difference, vertex$supports$Schaefer300,
  source_support=ngeo_vertex_area(vertex$difference)
)
budget <- ngeo_resource_budget(
  memory_bytes=8 * 1024^2,
  materialized_elements=300
)
regional <- aggregate_to(vertex$difference, map$target, map, budget=budget)
regional
```

```{r diagnostic, fig.width=7, fig.height=4, fig.alt="Sparse membership diagnostic for a 672 vertex by 300 parcel support map", fig.cap="The bounded operation consumes the sparse operator without materializing a dense whole-brain matrix."}
plot(map, type="membership", col="#2a9d8f", border="white")
```

If the declared limit is too small, the operation fails with a resource error
instead of switching atlas scale, dropping subjects, or choosing another
statistic. Record resource budgets with the analysis history and replay
manifest so execution decisions remain auditable.
