---
title: "Brain-GIS inference workflow in neurogeo 6.2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Brain-GIS inference workflow in neurogeo 6.2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Version 6.2 separates reference-map inference, partition training, and
population inference. These are different sampling units and must not be
reported as interchangeable tests.

## Local cross-layer coupling

Use Moran spectral randomization when the lagged reference maps' observed
global spatial autocorrelation should be retained. `maxT_scope = "subject"`
makes all requested local endpoints within each independent unit one max-T
family. `null = "free"` remains available only as a non-spatial reference-map
null under complete element exchangeability.

```{r, eval=FALSE}
local <- ngeo_local_layer_coupling(
  x, index, spatial_weights,
  metrics = "local_cross_moran",
  direction = "x_to_y",
  permutations = 999,
  null = "moran",
  adjust = "maxT",
  maxT_scope = "subject"
)
```

## Train, freeze, then apply a regionalization

Never learn a partition from the outcome data used for its downstream test.
Train on a separate dataset, freeze only its membership, and apply that frozen
membership to an independent dataset on the same spatial base.

```{r, eval=FALSE}
trained <- ngeo_contiguous_regionalization(
  training_data, spatial_weights,
  layers = c("thickness", "myelin"), n_regions = 10
)
frozen <- ngeo_freeze_regionalization(trained)
test_partition <- ngeo_apply_regionalization(test_data, frozen)
test_regions <- ngeo_aggregate(test_data, test_partition)
```

Exact reuse of the selected training values is rejected by default. Setting
`allow_training_data = TRUE` marks an application as descriptive and does not
make a training-set test valid.

## Point-process envelopes

`envelope = "pointwise"` controls each radius separately.
`envelope = "global"` uses a studentized maximum-deviation envelope across all
radii within each marked-pair family. `"both"` returns both summaries.
Unequal-exposure simple processes use exposure-weighted sampling without
replacement and require simulations for a sampling-design-matched CSR
normalization.

```{r, eval=FALSE}
pattern <- ngeo_brain_point_process(
  x, events, radii,
  exposure = opportunity,
  occupancy = "simple",
  simulations = 999,
  envelope = "global"
)
```

## Space-time hotspots and group inference

Hotspot inference jointly permutes spatial identities and assignments to the
declared real-time coordinates. Max-T controls the family of all local
space-time cells within one independent unit. The time bandwidth and cutoff
are expressed in `ngeo_time_axis()` units, not column indices.

```{r, eval=FALSE}
subject_hotspot <- ngeo_nonseparable_hotspots(
  subject_time_maps, spatial_weights,
  temporal_bandwidth = 7,
  temporal_cutoff = 21,
  permutations = 999,
  adjust = "maxT"
)

features <- ngeo_hotspot_group_features(
  list(subject_01 = subject_hotspot, subject_02 = subject_hotspot_02),
  summaries = c("mean_z", "trend")
)
group_result <- ngeo_group_test(
  features, subject_design, ~ age + group, test = "group",
  exchangeability = subject_exchangeability,
  adjustment = "maxT"
)
```

The joint hotspot null is still reference-map inference. The group test is the
population-level interface and requires one independently computed hotspot
result per declared sampling unit.

## Brain-landscape terminology

`ngeo_brain_landscape()` reports `boundary_edges` and
`support_per_boundary_edge_squared`. They are graph-topology summaries, not a
physical perimeter or a metric compactness measurement.
