---
title: "Change of support and cross-atlas analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Change of support and cross-atlas analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.retina = 2,
  out.width = "100%"
)
library(neurogeo)
```

## Background

Spatial support is the area, volume, vertex set, or other footprint over which
a value is defined. Changing support is therefore not a simple rename: it is
a linear or normalized transfer from source elements to target elements. Let
\(A\) be a target-by-source sparse operator. For extensive variables the
transfer is based on \(A x\); for intensive variables the weighted numerator
must be divided by transferred support.

`neurogeo` represents this relationship with one sparse operator whose rows
are target elements and columns are source elements.

## Build and audit a support map

```{r}
source <- ngeo_point(
  cbind(x = 1:4, y = 0),
  values = cbind(intensity = c(10, 20, 30, 40)),
  measures = ngeo_measure(support_behavior = "intensive")
)
target <- ngeo_parcellation(
  data.frame(region_id = c("A", "B")),
  support_size = c(3, 3)
)
support_map <- ngeo_support_map(
  source,
  target,
  operator = c("A", "A", "B", "B"),
  source_support = c(1, 2, 1, 2)
)
regional <- aggregate_to(source, target, support_map)
regional
```

The operator is easier to audit when the source elements and target summaries
are shown together. Point size below represents source support, not intensity.

```{r support-flow, fig.width=8, fig.height=4.5, fig.alt="Source points assigned to two colored regions alongside target regional intensity bars", fig.cap="A visual support audit separates source value, source support, and target membership."}
membership <- c("A", "A", "B", "B")
region_colors <- c(A = "#3b82a0", B = "#e29b45")
old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
plot(
  source$base$geometry$coordinates,
  type = "n",
  asp = 1,
  xlab = "x",
  ylab = "y",
  ylim = c(-0.5, 0.5),
  main = "Source support"
)
points(
  source$base$geometry$coordinates,
  pch = 21,
  bg = unname(region_colors[membership]),
  col = "white",
  cex = 1.5 + source$base$geometry$support_size
)
text(
  source$base$geometry$coordinates,
  labels = source$values[, "intensity"],
  pos = 3
)
barplot(
  regional$values[, "intensity"],
  names.arg = target$base$elements$region_id,
  col = unname(region_colors[target$base$elements$region_id]),
  border = NA,
  ylab = "Support-weighted intensity",
  main = "Target support"
)
par(old_par)
```

The bounded built-in diagnostic shows whether each source column has the
expected membership sum. It does not materialize or draw a whole-brain dense
operator.

```{r support-diagnostic, fig.alt="Histogram of source membership column sums", fig.cap="Every source element in this crisp complete map has membership sum one."}
plot(support_map, type = "membership", col = "#26a69a", border = "white")
```

## Intensive and extensive variables

Intensive layers use source-support weighted means. Extensive layers and counts
use conservative allocation. Overlapping operators therefore require
explicit column normalization before they can conserve totals.

## Compare atlases

Two atlas layers with a common source base can be compared by physical
support overlap:

```{r}
target2 <- ngeo_parcellation(
  data.frame(region_id = c("C", "D")),
  support_size = c(3, 3)
)
support_map2 <- ngeo_support_map(
  source,
  target2,
  operator = c("C", "D", "D", "C"),
  source_support = c(1, 2, 1, 2)
)
overlap <- as.matrix(
  ngeo_atlas_overlap(support_map, support_map2, distance_method = "dice")
)
overlap
```

```{r atlas-overlap, fig.width=6, fig.height=5, fig.alt="Heatmap of Dice overlap between two pairs of atlas regions", fig.cap="Cross-atlas overlap is a relationship between supports, not a claim that one atlas reconstructs the other."}
image(
  seq_len(nrow(overlap)),
  seq_len(ncol(overlap)),
  overlap,
  col = grDevices::hcl.colors(32, "Teal"),
  axes = FALSE,
  xlab = "Second atlas",
  ylab = "First atlas",
  main = "Dice overlap"
)
axis(1, at = seq_len(ncol(overlap)), labels = colnames(overlap))
axis(2, at = seq_len(nrow(overlap)), labels = rownames(overlap))
text(
  rep(seq_len(nrow(overlap)), times = ncol(overlap)),
  rep(seq_len(ncol(overlap)), each = nrow(overlap)),
  labels = sprintf("%.2f", as.vector(overlap))
)
```

`ngeo_cross_atlas()` is deliberately model based. It declares a
piecewise-constant source-parcel model and exposes the resulting transfer
operator and propagated variance. It is not described as an inverse
reconstruction.
