---
title: "Change of support: vertex SCZ-control maps across DK and Schaefer scales"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Change of support: vertex SCZ-control maps across DK and Schaefer scales}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
vertex <- ngeo_tutorial_vertex_case_control(n_per_group=5)
```

## Start from a vertex-level brain map

The sample contains five HC and five SCZ surface maps. The bilateral teaching
surface uses the 64,984-vertex Conte69 geometry and its registered flat chart.
The simulated SCZ effect is anterior-temporal and smooth.

```{r vertex-map, fig.width=9, fig.height=4.5, fig.alt="Vertex-level SCZ minus HC cortical thickness on a bilateral surface", fig.cap="Change of support starts from a spatially located vertex map, not an anonymous vector."}
vertex_values <- ngeo_values(vertex$difference)[, "SCZ_minus_HC"]
vertex_limit <- max(abs(vertex_values), na.rm=TRUE)
vertex_map <- ngeo_cortical_map(
  vertex$difference,
  layer="SCZ_minus_HC",
  chart="flat",
  atlas=vertex$supports$DK68,
  underlay=vertex$underlay,
  underlay_palette="Grays",
  overlay_alpha=0.82,
  palette="Blue-Red 3",
  limits=c(-vertex_limit, vertex_limit),
  na_color=NA_character_,
  atlas_coverage="auto"
)
plot(
  vertex_map,
  main="Vertex-level SCZ minus HC cortical thickness",
  boundary_color=grDevices::adjustcolor("white", 0.55),
  boundary_lwd=0.25,
  outline_lwd=1.1
)
```

The included DK68, Schaefer100/200/300, and Glasser360 memberships are the
published atlas labels registered to the same Conte69 vertex base.

```{r support-family}
vertex$support_note
vapply(vertex$supports, function(labels) length(unique(labels)), integer(1))
```

## Build sparse vertex-to-parcel operators

For cortical thickness, vertex area is the source support and the correct
parcel summary is an area-weighted mean. Each operator has target parcels in
rows and source vertices in columns.

```{r build-maps}
vertex_area <- ngeo_vertex_area(vertex$difference)
maps <- lapply(vertex$supports, function(labels) {
  ngeo_atlas_map(
    vertex$difference, labels,
    source_support=vertex_area
  )
})
diagnostics <- lapply(maps, ngeo_support_diagnostics)
data.frame(
  support=names(maps),
  targets=vapply(maps, function(map) nrow(map$operator), integer(1)),
  sources=vapply(maps, function(map) ncol(map$operator), integer(1)),
  coverage=vapply(maps, function(map) map$coverage, character(1))
)
```

```{r operator-diagnostic, fig.width=7, fig.height=4, fig.alt="Membership column sums for the vertex to DK68 support operator", fig.cap="Every included vertex contributes to exactly one parcel in a complete crisp atlas map."}
plot(maps$DK68, type="membership", col="#2a9d8f", border="white")
```

## Aggregate, then return the result to the brain

`aggregate_to()` uses the measure semantics: thickness is intensive, so the
operator transfers an area-weighted numerator and denominator. The parcel
result is lifted back to its source vertices only for visualization; it remains
a parcel-supported estimate.

```{r aggregate}
dk_result <- aggregate_to(vertex$difference, maps$DK68$target, maps$DK68)
dk_id <- ngeo_base_elements(dk_result)$region_id
dk_value <- ngeo_values(dk_result)[, "SCZ_minus_HC"]
dk_values <- setNames(dk_value, dk_id)
dk_limit <- max(abs(dk_values), na.rm=TRUE)
dk_map <- ngeo_cortical_map(
  vertex$difference,
  values=dk_values,
  chart="flat",
  atlas=vertex$supports$DK68,
  underlay=vertex$underlay,
  underlay_palette="Grays",
  overlay_alpha=0.82,
  palette="Blue-Red 3",
  limits=c(-dk_limit, dk_limit),
  na_color=NA_character_,
  atlas_coverage="auto"
)
```

```{r aggregate-brain, fig.width=9, fig.height=4.5, fig.alt="DK68 aggregated case-control thickness lifted to the bilateral surface", fig.cap="Piecewise-constant color makes the change from vertex support to parcel support visible."}
plot(
  dk_map,
  main="DK68-supported SCZ minus HC",
  boundary_color=grDevices::adjustcolor("white", 0.55),
  boundary_lwd=0.25,
  outline_lwd=1.1
)
```

The block-like map is not a reconstructed vertex effect. Every vertex in a DK
parcel is displaying the same parcel estimate.

## Compare declared atlas supports

```{r family-aggregation}
family_results <- lapply(maps, function(map) {
  aggregate_to(vertex$difference, map$target, map)
})
data.frame(
  support=names(family_results),
  n_parcel=vapply(family_results, function(x) nrow(ngeo_values(x)), integer(1)),
  contrast_sd=vapply(family_results, function(x) {
    stats::sd(ngeo_values(x)[, 1L])
  }, numeric(1))
)
```

```{r overlap, fig.width=8, fig.height=5, fig.alt="Dice overlap heatmap between synthetic DK68 and Schaefer100-sized supports", fig.cap="Atlas overlap is a relationship between supports, not evidence that one atlas reconstructs the other."}
overlap <- as.matrix(ngeo_atlas_overlap(
  maps$DK68, maps$Schaefer100, distance_method="dice"
))
image(t(overlap[nrow(overlap):1, ]), axes=FALSE,
      col=grDevices::hcl.colors(40, "Teal"),
      xlab="Schaefer100-sized parcels", ylab="DK68-sized parcels",
      main="Cross-support Dice overlap")
```

Cross-atlas agreement must be evaluated over the declared support family. A
result consistent across these registered atlas scales still requires the same
subject-level design on every support before it can support a robustness claim.
