---
title: "Change of support and cross-atlas analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Change of support and cross-atlas analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

NGCS 2.0 represents support change with one sparse operator whose rows are
target elements and columns are source elements.

```{r}
source <- ngeo_points(
  cbind(x = 1:4, y = 0),
  values = cbind(intensity = c(10, 20, 30, 40)),
  measures = ngeo_measure(spatial_semantics = "intensive")
)
target <- ngeo_regions(
  data.frame(region_id = c("A", "B")),
  support_size = c(3, 3)
)
support_map <- ngeo_support_map(
  source,
  target,
  operator = c("A", "A", "B", "B"),
  source_support = c(1, 2, 1, 2)
)
ngeo_change_support(source, target, support_map)
```

Intensive maps use source-support weighted means. Extensive maps and counts
use conservative allocation. Overlapping operators therefore require
explicit column normalization before they can conserve totals.

Two atlas maps with a common source domain can be compared by physical
support overlap:

```{r}
target2 <- ngeo_regions(
  data.frame(region_id = c("C", "D")),
  support_size = c(3, 3)
)
support_map2 <- ngeo_support_map(
  source,
  target2,
  operator = c("C", "D", "D", "C"),
  source_support = c(1, 2, 1, 2)
)
as.matrix(ngeo_atlas_overlap(support_map, support_map2, metric = "dice"))
```

`ngeo_cross_atlas()` is deliberately model based. It declares a
piecewise-constant source-parcel model and exposes the resulting transfer
operator and propagated variance. It is not described as an inverse
reconstruction.
