---
title: "Core concepts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core concepts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

`neurogeo` represents one spatial domain and one strictly aligned values
block. Space, topology, metric, measurement semantics, and provenance remain
explicit instead of being inferred from filenames or column names.

```{r}
coordinates <- matrix(
  c(0, 0, 0, 1, 0, 0, 0, 1, 0),
  ncol = 3,
  byrow = TRUE
)
surface <- ngeo_surface(
  coordinates,
  matrix(c(1, 2, 3), nrow = 1),
  values = cbind(thickness = c(2.1, 2.4, 2.2)),
  measures = ngeo_measure(
    spatial_semantics = "intensive",
    units = "mm"
  )
)
surface
ngeo_capabilities(surface)
ngeo_domain_hash(surface)
```

Rows of `values` are aligned to `ngeo_elements(surface)` and columns are
described by both `ngeo_maps(surface)` and `ngeo_measures(surface)`. A subset
operation updates all aligned structures and records provenance.

```{r}
subset <- ngeo_subset(surface, elements = c(1, 3))
ngeo_elements(subset)
ngeo_provenance(subset)$operations
```

Topology and metric are distinct. Adjacency says which vertices share an
edge; edge-geodesic distance assigns lengths to those edges and finds a
shortest path.

```{r}
ngeo_adjacency(surface)
ngeo_distance(surface, from = 1, to = 2, metric = "edge_geodesic")
```
