---
title: "Core concepts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core concepts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Neuroimaging measurements are meaningful only when their values remain tied
to the spatial elements on which they were observed. A row may describe a
surface vertex, voxel, point, grayordinate, or region; changing that order
without changing the values changes the data.

`neurogeo` therefore represents one spatial domain and one strictly aligned
values block. Space, topology, metric, measurement semantics, and provenance
remain explicit instead of being inferred from filenames or column names.
This contract separates *where an observation is defined* from *what was
measured there*.

## Domain and aligned values

```{r}
coordinates <- matrix(
  c(0, 0, 0, 1, 0, 0, 0, 1, 0),
  ncol = 3,
  byrow = TRUE
)
surface <- ngeo_surface(
  coordinates,
  matrix(c(1, 2, 3), nrow = 1),
  values = cbind(thickness = c(2.1, 2.4, 2.2)),
  measures = ngeo_measure(
    spatial_semantics = "intensive",
    units = "mm"
  )
)
surface
ngeo_capabilities(surface)
ngeo_domain_hash(surface)
```

Rows of `values` are aligned to `ngeo_elements(surface)` and columns are
described by both `ngeo_maps(surface)` and `ngeo_measures(surface)`. A subset
operation updates all aligned structures and records provenance.

```{r}
subset <- ngeo_subset(surface, elements = c(1, 3))
ngeo_elements(subset)
ngeo_provenance(subset)$operations
```

## Topology and metric

Topology and metric answer different questions. Topology records which
elements are connected; a metric assigns a distance to pairs or paths. Two
vertices may be adjacent in a mesh even though the physical length of their
connecting edge differs from other edges. Edge-geodesic distance assigns
lengths to those edges and finds a shortest path.

```{r}
ngeo_adjacency(surface)
ngeo_distance(surface, from = 1, to = 2, metric = "edge_geodesic")
```
