---
title: "Core concepts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core concepts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Neuroimaging measurements are meaningful only when their values remain tied
to the spatial elements on which they were observed. A row may describe a
surface vertex, voxel, point, grayordinate, or region; changing that order
without changing the values changes the data.

`neurogeo` therefore represents one spatial base and one strictly aligned
values block. Space, topology, distance_method, measurement semantics, and history
remain explicit instead of being inferred from filenames or column names.
This contract separates *where an observation is defined* from *what was
measured there*.

## Domain and aligned values

```{r}
coordinates <- matrix(
  c(0, 0, 0, 1, 0, 0, 0, 1, 0),
  ncol = 3,
  byrow = TRUE
)
surface <- ngeo_surface(
  coordinates,
  matrix(c(1, 2, 3), nrow = 1),
  values = cbind(thickness = c(2.1, 2.4, 2.2)),
  measures = ngeo_measure(
    support_behavior = "intensive",
    unit = "mm"
  )
)
surface
ngeo_capabilities(surface)
base_hash(surface)
```

Rows of `values` are aligned to `base_elements(surface)` and columns are
described by both `layers(surface)` and `measures(surface)`. A subset
operation updates all aligned structures and records history.

```{r}
subset <- ngeo_subset(surface, elements = c(1, 3))
base_elements(subset)
history(subset)$operations
```

## Topology and distance_method

Topology and distance_method answer different questions. Topology records which
elements are connected; a distance_method assigns a distance to pairs or paths. Two
vertices may be adjacent in a mesh even though the physical length of their
connecting edge differs from other edges. Edge-geodesic distance assigns
lengths to those edges and finds a shortest path.

```{r}
ngeo_adjacency(surface)
ngeo_distance(surface, from = 1, to = 2, distance_method = "edge_geodesic")
```
