---
title: "Cortical cartography for arbitrary vertex data and atlases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cortical cartography for arbitrary vertex data and atlases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## One source surface, no fixed atlas drawing

The 4.3 cartography module uses the geometry and topology of the actual
`ngeo_surface`. Atlas names never select hidden template geometry. A chart,
source vertices, source faces, distortion, space, and provenance remain
inspectable.

```{r disk}
coordinates <- matrix(
  c(
    0, 0, 0, 1, 0, 0, 1, 1, 0, 0, 1, 0,
    0.5, 0.5, 0.2
  ),
  ncol = 3,
  byrow = TRUE
)
faces <- matrix(
  c(1, 2, 5, 2, 3, 5, 3, 4, 5, 4, 1, 5),
  ncol = 3,
  byrow = TRUE
)
surface <- ngeo_surface(
  coordinates,
  faces,
  values = cbind(thickness = c(2.1, 2.4, 2.9, 2.5, 2.7)),
  measures = ngeo_measure(
    value_type = "continuous",
    spatial_semantics = "intensive",
    units = "mm"
  )
)
atlas_labels <- c(
  "frontal", "frontal", "parietal", "parietal", "central"
)
atlas <- ngeo_partition(surface, atlas_labels)
```

## Parameterize an explicit disk

Harmonic parameterization requires one connected triangulated disk and its
complete ordered boundary. It never cuts a closed surface.

```{r flatten}
flat <- ngeo_flatten_surface(
  surface,
  "harmonic",
  boundary = c(1, 2, 3, 4)
)
flat$domain$charts$flat$distortion_summary
head(ngeo_chart_distortion(flat, "flat"))
```

A caller-supplied flat coordinate matrix can instead be preserved with
`method = "imported"`.

## Map aligned vertex values and any aligned atlas

```{r map, fig.width=5, fig.height=5}
map <- ngeo_cortical_map(
  flat,
  map = "thickness",
  atlas = atlas
)
plot(map)
```

An arbitrary aligned atlas can itself be rendered as categorical vertex
data without substituting another geometry. Multiple maps or hemispheres
remain independent when arranged:

```{r layout, fig.width=8, fig.height=4}
atlas_map <- ngeo_cortical_map(
  flat,
  values = atlas_labels,
  atlas = atlas,
  palette = "Set 3"
)
plot(ngeo_cortical_layout(map, atlas_map, ncol = 2))
```

## Exchange plotting data

```{r exchange}
data <- ngeo_cortical_map_data(map)
names(data)
head(data$vertices)
head(data$faces)
head(data$boundaries)
```

Downstream renderers should preserve source indices, chart metadata, and
seam metadata.

## Viewing projections are not flattening

```{r view}
viewed <- ngeo_project_surface(surface, "pca")
viewed$domain$charts$view$kind
viewed$domain$charts$view$is_metric_flattening
```

Orthographic, PCA, and spherical outputs are explicitly non-metric views.
Spherical viewing requires a seam longitude and records seam-crossing faces.
The module does not reconstruct surfaces, estimate registration, resample
data, perform inference, or run quality control.
