---
title: "Real cortical flatmaps for vertex data and arbitrary atlases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real cortical flatmaps for vertex data and arbitrary atlases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
flatmap_figure <- function(name, alt) {
  if (identical(Sys.getenv("NEUROGEO_VITEPRESS"), "true")) {
    knitr::asis_output(sprintf(
      '<img src="/images/%s" alt="%s" style="width: 100%%;" />',
      name,
      alt
    ))
  } else {
    knitr::include_graphics(file.path("figures", name))
  }
}
```

## A real cortical sheet, not a template cartoon

This module renders a registered cortical flat surface with its brain-shaped
outline, anatomical underlay, vertex data, atlas boundaries, and labels. It
does not substitute a fixed atlas drawing or a toy polygon for the caller's
surface.

The following release-validation figures use the HCP S1200 Conte69 32k flat
surfaces. The first overlays a continuous vertex metric. The second overlays
the Schaefer 2018 atlas aggregated to seven networks. A geometry-derived
curvature diagnostic supplies the grayscale folding underlay.

```{r real-continuous-figure, echo=FALSE, results="asis", out.width="100%"}
flatmap_figure(
  "conte69-vertex-flatmap.png",
  "Bilateral Conte69 flat surfaces with continuous vertex data, an anatomical underlay, atlas boundaries, and cortical outlines."
)
```

*Bilateral Conte69 flat surfaces with continuous vertex data, an anatomical
underlay, atlas boundaries, and cortical outlines.*

```{r real-atlas-figure, echo=FALSE, results="asis", out.width="100%"}
flatmap_figure(
  "conte69-atlas-flatmap.png",
  "The same real geometry with the Schaefer atlas aggregated to seven networks, including boundaries and labels."
)
```

*The same real geometry with the Schaefer atlas aggregated to seven networks,
including boundaries and labels.*

## Data contract

A defensible map keeps five aligned inputs separate:

1. a 3D source surface and topology;
2. a registered flat surface;
3. a cortical mask;
4. an anatomical underlay such as sulcal depth or curvature;
5. the vertex values or atlas membership to display.

An imported flat surface is verified against ordered source vertices. Every
flat triangle must map to a source face. A verified face subset is allowed
because a real cortical cut or medial-wall treatment can remove triangles;
an invented or unmapped triangle fails.

## Bind an existing registered flat surface

Replace these paths with aligned GIFTI files from the same surface system:

```{r read-real-surface, eval=FALSE}
source_surface <- read_ngeo_gifti(
  geometry = "S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii",
  data = c(
    metric = "metric.shape.gii",
    sulc = "sulc.shape.gii",
    atlasroi = "L.atlasroi.32k_fs_LR.shape.gii"
  ),
  labels = c(
    schaefer100 = "schaefer-100_conte69_lh.label.gii"
  ),
  checksum = TRUE
)

flat_surface <- read_ngeo_gifti(
  "S1200.L.flat.32k_fs_LR.surf.gii",
  checksum = TRUE
)

left <- ngeo_flatten_surface(
  source_surface,
  method = "imported",
  coordinates = flat_surface,
  name = "flat"
)

left$domain$charts$flat$invariants[
  c("topology_verified", "topology_relation")
]
```

This operation does not infer a cut, registration, or resampling. Misaligned
surfaces fail instead of producing an unauditable picture.

## Continuous vertex map

```{r continuous-map, eval=FALSE}
continuous_map <- ngeo_cortical_map(
  left,
  map = "metric",
  chart = "flat",
  atlas = "schaefer100",
  mask = left$values[, "atlasroi"] > 0,
  underlay = "sulc",
  underlay_palette = "Grays",
  overlay_alpha = 0.78,
  palette = "viridis",
  na_color = NA_character_
)

plot(
  continuous_map,
  show_boundaries = TRUE,
  boundary_color = adjustcolor("white", 0.4),
  boundary_lwd = 0.2,
  show_outline = TRUE,
  outline_lwd = 1.4
)
```

Continuous face colors are finite vertex means, while the original aligned
vertex values remain in the exchange object.

## Any aligned atlas

The atlas may be an aligned atomic vector, an `ngeo_partition`, or the name
of an aligned label map read from GIFTI. Brodmann, Schaefer, Glasser,
visuotopic, and study-specific atlases therefore use the same rendering
path; the atlas never replaces surface geometry.

```{r atlas-map, eval=FALSE}
atlas_map <- ngeo_cortical_map(
  left,
  chart = "flat",
  atlas = "schaefer100",
  fill = "atlas",
  mask = left$values[, "atlasroi"] > 0,
  underlay = "sulc",
  underlay_palette = "Grays",
  overlay_alpha = 0.82,
  na_color = NA_character_
)

plot(
  atlas_map,
  show_boundaries = TRUE,
  boundary_lwd = 0.55,
  show_labels = TRUE,
  label_cex = 0.45,
  show_outline = TRUE
)
```

GIFTI label-table names and RGBA colors are reused automatically for
`fill = "atlas"`. Named colors can override them. Setting non-target atlas
membership to `NA` creates a partial overlay while preserving the complete
grayscale cortical sheet underneath.

## Bilateral layouts and exchange data

```{r bilateral-and-exchange, eval=FALSE}
plot(ngeo_cortical_layout(
  left_map,
  right_map,
  ncol = 2,
  labels = c("Left hemisphere", "Right hemisphere"),
  shared_scale = TRUE
))

plot_data <- ngeo_cortical_map_data(atlas_map)
names(plot_data)
# vertices, faces, boundaries, outline, label_positions, legend, metadata

plot_data$metadata$chart_metadata$invariants$source_face_in_chart
plot_data$metadata$provenance
```

Panels retain independent domains. Downstream renderers should preserve
source vertex/face indices, inclusion status, chart metadata, and
provenance.

## Scientific boundary

The module does not reconstruct surfaces, cut a closed hemisphere,
register/resample data, perform inference, or run QC. Orthographic and PCA
outputs are views, not cortical flattening. A formal result should report
the source surface, flat-chart source, topology verification, mask,
underlay, atlas, color limits, and provenance.

The validation fixtures are governed by the
[HCP Open Access Data Use Terms](https://www.humanconnectome.org/study/hcp-young-adult/document/wu-minn-hcp-consortium-open-access-data-use-terms).
The Schaefer atlas should be cited as Schaefer et al. (2018),
*Cerebral Cortex*, 28(9):3095–3114. Checksums and immutable sources are
recorded in `inst/extdata/reference-4.3.1/manifest.csv`; the neuroimaging
files are not bundled with neurogeo.
