---
title: "File-backed neuroimaging values"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{File-backed neuroimaging values}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Background

File-backed access delays reading until a particular element-by-map block is
requested. It is useful for long time series and dense grayordinate data, but
it is not a different scientific data model. The file identity and selection
must be part of the object identity so that a changed source file cannot be
mistaken for the same data.

The object keeps one spatial domain and one aligned element-by-map values
block, but the block can address a source file instead of retaining the
complete matrix.

## File-backed readers

```{r}
library(neurogeo)

x <- read_ngeo_nifti_filebacked(
  "sub-01_task-rest_bold.nii.gz",
  frames = 1:20,
  verify = "checksum",
  budget = ngeo_resource_budget(
    memory_bytes = 64 * 1024^2,
    materialized_elements = 8e6
  )
)

x$values[1:1000, 1:2, drop = FALSE]
```

The same contract is available for CIFTI and FreeSurfer volumes.

```{r}
dense <- read_ngeo_cifti_filebacked(
  "sub-01_task-rest_bold.dtseries.nii",
  structures = c("CORTEX_LEFT", "CORTEX_RIGHT"),
  frames = 1:100
)

mgh <- read_ngeo_mgh_filebacked("lh.metric.mgz", frames = 1)
```

## Selection and identity

Selections are reflected in both the domain and map table. A file identity
binds the exact selection and source state, so it is safe to include in cache
and checkpoint identities.

## Bounded streaming

Use deterministic chunks for streaming methods.

```{r}
summaries <- ngeo_value_chunks(
  dense,
  chunk_size = 4096,
  FUN = function(block, rows) {
    list(rows = range(rows), mean = colMeans(block))
  }
)
```

The resource budget is checked for each requested block. Requesting a block
larger than the budget raises `ngeo_error_resource` before file data are
read.

## Writing

`write_ngeo_filebacked()` is intentionally narrow: it atomically copies a
complete source through a bounded buffer. It rejects partial selections,
because copying the original bytes would otherwise publish data that are not
represented by the selected object.

```{r}
complete <- read_ngeo_filebacked("image.nii.gz")
write_ngeo_filebacked(complete, "image-copy.nii.gz")
```

Use the normal format writer to encode a deliberately materialized subset.
Neither file-backed readers nor writers perform registration, resampling, or
preprocessing.
