---
title: "Choose a workflow: NIfTI, surface, CIFTI, or ROI/cohort"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choose a workflow: NIfTI, surface, CIFTI, or ROI/cohort}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Choose from the spatial element—not the filename alone

Use this page after the [15-minute quickstart](/en/tutorials/getting-started). A filename tells you its container format; the analysis base tells you what one row means.

| Your rows represent | Typical input | Base | First checks |
|---|---|---|---|
| voxels | NIfTI | volume | dimensions, affine, mask, physical unit |
| surface vertices | GIFTI or FreeSurfer | surface | vertex order, faces, coordinate role |
| mixed brain models | CIFTI | grayordinate | component order, structure, vertex/voxel index |
| parcels × subjects | CSV/TSV | parcellation plus subject layers | region IDs, subject IDs, measure semantics |

The installed teaching corpus has one CC0 fixture for each route:

```{r corpus}
corpus_path <- system.file(
  "extdata", "tutorial-fixtures-6.0.csv",
  package = "neurogeo",
  mustWork = TRUE
)
corpus <- read.csv(corpus_path)
corpus[c("fixture_id", "format", "workflow", "size_bytes", "sha256")]
```

## NIfTI / volume

Start with the affine, not a screenshot. Equal array dimensions do not imply the same coordinate space.

```{r nifti-path}
nifti_path <- system.file(
  "extdata", "golden", "tiny.nii.gz",
  package = "neurogeo", mustWork = TRUE
)
nifti_path
```

Continue with the executable [NIfTI volume workflow](/en/tutorials/workflow-volume). Report the chosen world-space distance and verified spatial unit.

## GIFTI or FreeSurfer / surface

A surface needs both ordered vertex coordinates and triangular faces. Anatomical coordinates may define a metric; registration and visualization coordinates must not silently do so.

```{r surface-path}
surface_path <- system.file(
  "extdata", "golden", "tetra.surf.gii",
  package = "neurogeo", mustWork = TRUE
)
surface_path
```

Continue with the executable [surface workflow](/en/tutorials/workflow-surface) after verifying faces, components, coordinate roles, and masks.

## CIFTI / grayordinate

A grayordinate row belongs to an ordered brain model. It may refer to a cortical vertex or a subcortical voxel; there is no valid implicit distance across structures.

```{r cifti-path}
cifti_path <- system.file(
  "extdata", "golden", "tiny.dscalar.nii",
  package = "neurogeo", mustWork = TRUE
)
cifti_path
```

Check every component's structure and source indices. Surface distance also requires matching surface geometry; the CIFTI scalar array alone does not provide it.

Continue with the executable [CIFTI workflow](/en/tutorials/workflow-cifti).

## ROI × cohort

Keep region support separate from independent subjects. Regions define the spatial base; subjects normally belong in the layer table or group-design object rather than being treated as spatial neighbors.

```{r roi-path}
roi_path <- system.file(
  "extdata", "golden", "tiny-roi-cohort.csv",
  package = "neurogeo", mustWork = TRUE
)
roi_cohort <- read.csv(roi_path)
roi_cohort
```

Before group inference, declare the independent unit and exchangeability schedule. Do not treat several atlases derived from the same subjects as independent studies.

Continue with the executable [ROI-by-cohort workflow](/en/tutorials/workflow-roi-cohort).

## What all four routes share

Every workflow should reach the same checkpoints:

1. read from a real immutable source;
2. inspect ordered elements and values;
3. confirm coordinate space and measurement semantics;
4. run strict validation and QC;
5. declare metric, support, and null explicitly;
6. inspect `ngeo_inference_contract()` for a scientific result;
7. write output with history and source identity.

The format-specific end-to-end tutorials expand these checkpoints without changing the core object model.
