---
title: "15-minute quickstart: a DK68 SCZ-control cortical study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{15-minute quickstart: a DK68 SCZ-control cortical study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE,
                      warning = FALSE, fig.align = "center")
library(neurogeo)
source(system.file(
  "tutorial-code", "brain-case-study.R", package = "neurogeo",
  mustWork = TRUE
))
dk <- ngeo_tutorial_dk_case_control(n_per_group = 100, seed = 20260810)
```

## The question, before the object

We simulate cortical thickness for 100 healthy controls (HC) and 100 people
with schizophrenia (SCZ), measured in the 68 cortical Desikan--Killiany (DK)
parcels. Age, sex, and site are subject-level covariates. The simulation embeds
lower SCZ thickness in bilateral superior temporal, insular, anterior
cingulate, and medial orbitofrontal cortex, with weaker effects in neighboring
parcels. It is teaching data, not clinical evidence.

```{r study-design}
with(dk$design, table(group, site))
summary(dk$design$age)
dim(ngeo_values(dk$cohort))
```

The independent sampling unit is the subject. DK parcels are spatial elements;
they are not 68 additional subjects.

## See the brain before computing a statistic

The first map is one HC participant. The second is the descriptive SCZ-minus-HC
contrast. Every subsequent number in this tutorial refers back to these maps.

```{r subject-brain, fig.width=9, fig.height=4.6, fig.alt="DK cortical thickness for one simulated healthy control shown on bilateral cortical views", fig.cap="One subject is one layer; DK parcels form the shared spatial base."}
ngeo_tutorial_plot_dk(
  dk$cohort, layer = "sub-001",
  title = "Simulated cortical thickness: sub-001 (HC)"
)
```

```{r contrast-brain, fig.width=9, fig.height=4.6, fig.alt="Simulated SCZ minus HC cortical thickness difference on the DK atlas", fig.cap="The case-control contrast is spatially structured and is plotted before spatial inference."}
ngeo_tutorial_plot_dk(
  dk$difference,
  title = "SCZ minus HC cortical thickness (mm)",
  midpoint = 0
)
```

## Data model: base, values, layers, and measure

| Term | Neuroimaging meaning here | Shape |
|---|---|---|
| base | DK cortical parcels | 68 rows |
| values | cortical thickness | 68 × 200 |
| layer | one subject map | 200 columns |
| measure | thickness, intensive, mm | shared semantics |

```{r inspect}
ngeo_base_type(dk$cohort)
head(ngeo_base_elements(dk$cohort)[c("region_id", "hemi", "region", "lobe")])
head(ngeo_layers(dk$cohort)[c("layer_id", "subject_id", "group", "age", "site")])
ngeo_measures(dk$cohort)
ngeo_validate(dk$cohort, "strict")
```

Row *i* always means the same DK parcel across all 200 columns. Subject
metadata belongs to layers; millimetres and intensive support semantics belong
to the measure. This alignment is the core contract that a numeric matrix alone
does not carry.

## Declare a cortical spatial relation

The teaching fixture includes a fixed symmetric DK neighborhood graph derived
from the atlas drawing. `region_contiguity` uses that graph directly. For a
real study, build adjacency or geodesic weights from the registered cortical
surface and record that provenance; do not infer scientific adjacency from a
figure.

```{r weights}
weights <- ngeo_spatial_weights(
  dk$difference,
  method = "region_contiguity",
  style = "W"
)
weights
summary(Matrix::rowSums(weights$raw_matrix != 0))
```

## Ask a spatial question about the group contrast

Global Moran's I below asks whether the *descriptive group-difference map* is
more spatially clustered than expected under value-label permutation on the
fixed DK graph. It does **not** test whether SCZ and HC differ: that requires
subject-level inference with the 200 independent subjects.

```{r moran}
spatial_result <- ngeo_moran(
  dk$difference,
  weights,
  layer = "SCZ_minus_HC",
  permutations = 499,
  seed = 2026
)
spatial_result
ngeo_inference_contract(spatial_result)
```

The contract makes the distinction auditable: its sampling units are parcels
under a spatial randomization null. In a case-control model, the sampling units
are subjects and the exchangeability rule acts on subjects, never vertices or
parcels.

## Localize the descriptive pattern

```{r local}
local <- ngeo_local_moran(
  dk$difference, weights, layer = "SCZ_minus_HC",
  permutations = 499, seed = 2026, adjust = "BH",
  null_model = "conditional"
)
head(local[order(local$p.adjusted),
  c("element_id", "local_i", "p.adjusted", "cluster")], 10)
```

Local Moran quadrants describe where similar contrast values cluster; adjusted
permutation evidence is required before calling a local pattern significant.
The brain map remains the primary orientation, while this table supplies exact
parcel-level values.

## What to report

> We simulated DK68 cortical thickness in 100 HC and 100 SCZ participants. A
> descriptive SCZ-minus-HC map was evaluated on a declared row-standardized DK
> adjacency graph using 499 value-label permutations. This spatial test
> characterizes clustering of the contrast map; it is not a subject-level test
> of diagnosis. Population inference must preserve subjects as the independent
> sampling unit and adjust the prespecified covariates.

Continue with [ROI/cohort data and I/O](/en/tutorials/workflow-roi-cohort),
[neighbors and weights](/en/tutorials/neighbors-and-weights),
[change of support](/en/tutorials/change-of-support), or
[multilayer subject-level inference](/en/modules/multilayer-inference).
