---
title: "15-minute quickstart: a DK68 SCZ-control cortical study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{15-minute quickstart: a DK68 SCZ-control cortical study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE,
                      warning = FALSE, fig.align = "center")
library(neurogeo)
source(system.file(
  "tutorial-code", "brain-case-study.R", package = "neurogeo",
  mustWork = TRUE
))
dk <- ngeo_tutorial_dk_case_control(n_per_group = 100, seed = 20260810)
vertex <- ngeo_tutorial_vertex_case_control(n_per_group = 5)
```

## The question, before the object

We simulate cortical thickness for 100 healthy controls (HC) and 100 people
with schizophrenia (SCZ), measured in the 68 cortical Desikan--Killiany (DK)
parcels. Age, sex, and site are subject-level covariates. The simulation embeds
lower SCZ thickness in bilateral superior temporal, insular, anterior
cingulate, and medial orbitofrontal cortex, with weaker effects in neighboring
parcels and a small smooth anterior--posterior background gradient. It is
teaching data, not clinical evidence.

```{r study-design}
with(dk$design, table(group, site))
summary(dk$design$age)
dim(ngeo_values(dk$cohort))
```

The independent sampling unit is the subject. DK parcels are spatial elements;
they are not 68 additional subjects.

## See the brain before computing a statistic

The first map is one HC participant. The second is the descriptive SCZ-minus-HC
contrast. Every subsequent number in this tutorial refers back to these maps.

```{r subject-brain, fig.width=9, fig.height=4.6, fig.alt="DK cortical thickness for one simulated healthy control shown on bilateral cortical views", fig.cap="One subject is one layer; DK parcels form the shared spatial base."}
subject_values <- setNames(
  ngeo_values(dk$cohort)[, "sub-001"],
  ngeo_base_elements(dk$cohort)$region_id
)
subject_map <- ngeo_cortical_map(
  dk$surface,
  values = subject_values,
  chart = "flat",
  atlas = dk$atlas,
  underlay = dk$underlay,
  underlay_palette = "Grays",
  overlay_alpha = 0.82,
  palette = "viridis",
  na_color = NA_character_,
  atlas_coverage = "auto"
)
plot(
  subject_map,
  main = "sub-001 cortical thickness (HC)",
  boundary_color = grDevices::adjustcolor("white", 0.55),
  boundary_lwd = 0.25,
  outline_lwd = 1.1
)
```

```{r contrast-brain, fig.width=9, fig.height=4.6, fig.alt="Simulated SCZ minus HC cortical thickness difference on the DK atlas", fig.cap="The case-control contrast is spatially structured and is plotted before spatial inference."}
contrast_values <- setNames(
  ngeo_values(dk$difference)[, "SCZ_minus_HC"],
  ngeo_base_elements(dk$difference)$region_id
)
contrast_limit <- max(abs(contrast_values), na.rm = TRUE)
contrast_map <- ngeo_cortical_map(
  dk$surface,
  values = contrast_values,
  chart = "flat",
  atlas = dk$atlas,
  underlay = dk$underlay,
  underlay_palette = "Grays",
  overlay_alpha = 0.82,
  palette = "Blue-Red 3",
  limits = c(-contrast_limit, contrast_limit),
  na_color = NA_character_,
  atlas_coverage = "auto"
)
plot(
  contrast_map,
  main = "SCZ - HC cortical thickness (mm)",
  boundary_color = grDevices::adjustcolor("white", 0.55),
  boundary_lwd = 0.25,
  outline_lwd = 1.1
)
```

## Data model: base, values, layers, and measure

| Term | Neuroimaging meaning here | Shape |
|---|---|---|
| base | DK cortical parcels | 68 rows |
| values | cortical thickness | 68 × 200 |
| layer | one subject map | 200 columns |
| measure | thickness, intensive, mm | shared semantics |

```{r inspect}
ngeo_base_type(dk$cohort)
head(ngeo_base_elements(dk$cohort)[c("region_id", "hemi", "region", "lobe")])
head(ngeo_layers(dk$cohort)[c("layer_id", "subject_id", "group", "age", "site")])
ngeo_measures(dk$cohort)
ngeo_validate(dk$cohort, "strict")
```

Row *i* always means the same DK parcel across all 200 columns. Subject
metadata belongs to layers; millimetres and intensive support semantics belong
to the measure. This alignment is the core contract that a numeric matrix alone
does not carry.

## Declare a cortical spatial relation

The teaching fixture includes a fixed symmetric DK neighborhood graph derived
from the atlas drawing. `region_contiguity` uses that graph directly. For a
real study, build adjacency or geodesic weights from the registered cortical
surface and record that provenance; do not infer scientific adjacency from a
figure.

```{r weights}
weights <- ngeo_spatial_weights(
  dk$difference,
  method = "region_contiguity",
  style = "W"
)
weights
summary(Matrix::rowSums(weights$raw_matrix != 0))
```

The same vertex base also carries several published atlas label sets.

```{r atlas-family, fig.width=10, fig.height=8, fig.alt="DK68 Schaefer100 Schaefer200 Schaefer300 and Glasser360 on Conte69 flat cortex", fig.cap="All atlases are strictly aligned to the same 64,984-vertex Conte69 base."}
atlas_names <- c(
  "DK68", "Schaefer100", "Schaefer200", "Schaefer300", "Glasser360"
)
atlas_maps <- lapply(atlas_names, function(atlas_name) {
  ngeo_cortical_map(
    vertex$surface,
    chart = "flat",
    atlas = vertex$supports[[atlas_name]],
    fill = "atlas",
    underlay = vertex$underlay,
    underlay_palette = "Grays",
    overlay_alpha = 0.78,
    palette = "Dynamic",
    na_color = NA_character_,
    atlas_coverage = "auto"
  )
})
atlas_layout <- ngeo_cortical_layout(
  atlas_maps,
  ncol = 2L,
  labels = atlas_names,
  shared_scale = FALSE
)
plot(
  atlas_layout,
  show_legend = FALSE,
  boundary_lwd = 0.35,
  outline_lwd = 1.1
)
```

## Ask a spatial question about the group contrast

Global Moran's I below asks whether the *descriptive group-difference map* is
more spatially clustered than expected under value-label permutation on the
fixed DK graph. It does **not** test whether SCZ and HC differ: that requires
subject-level inference with the 200 independent subjects.

```{r moran}
spatial_result <- ngeo_moran(
  dk$difference,
  weights,
  layer = "SCZ_minus_HC",
  permutations = 499,
  seed = 2026
)
spatial_result
contract <- ngeo_inference_contract(spatial_result)
contract
```

The contract makes the distinction auditable: its sampling units are parcels
under a spatial randomization null. In a case-control model, the sampling units
are subjects and the exchangeability rule acts on subjects, never vertices or
parcels.

## Localize the descriptive pattern

```{r local}
local <- ngeo_local_moran(
  dk$difference, weights, layer = "SCZ_minus_HC",
  permutations = 499, seed = 2026, adjust = "BH",
  null_model = "conditional"
)
head(local[order(local$p.adjusted),
  c("element_id", "local_i", "p.adjusted", "cluster")], 10)
```

Local Moran quadrants describe where similar contrast values cluster; adjusted
permutation evidence is required before calling a local pattern significant.
The brain map remains the primary orientation, while this table supplies exact
parcel-level values.

## What to report

> We simulated DK68 cortical thickness in 100 HC and 100 SCZ participants. A
> descriptive SCZ-minus-HC map was evaluated on a declared row-standardized DK
> adjacency graph using 499 value-label permutations. This spatial test
> characterizes clustering of the contrast map; it is not a subject-level test
> of diagnosis. Population inference must preserve subjects as the independent
> sampling unit and adjust the prespecified covariates.

Continue with [ROI/cohort data and I/O](/en/tutorials/workflow-roi-cohort),
[neighbors and weights](/en/tutorials/neighbors-and-weights),
[change of support](/en/tutorials/change-of-support), or
[multilayer subject-level inference](/en/modules/multilayer-inference).
