---
title: "15-minute quickstart: from a fixed fixture to interpretable Moran's I"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{15-minute quickstart: from a fixed fixture to interpretable Moran's I}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Who is this for?

Start here if you know MRI, fMRI, cortical surfaces, or ROI tables but are new to `neurogeo`. The tutorial takes about 10–15 minutes and less than 50 MB of memory. It runs in a clean R session with data installed in the package—there are no placeholder paths or downloads.

By the end, you will be able to:

1. distinguish a `base`, `values`, a `layer`, and a `measure`;
2. construct and validate a spatial dataset;
3. declare neighbors and spatial weights explicitly;
4. run and interpret global Moran's I;
5. inspect the null, metric, and sampling unit of the result.

## Four terms first

A neuroimaging array alone does not say where each row lives or whether values may be summed. `neurogeo` keeps four kinds of information separate:

| Term | Question | This example |
|---|---|---|
| spatial base | Where does each value row live? | 9 two-dimensional points |
| values | What was measured? | one `signal` column |
| layer | Which values column is analysed? | `signal` |
| measure | What are its unit and support semantics? | a.u. and intensive |

`intensive` means aggregation should use a support-weighted mean. It differs from `extensive` or `count` values that may be summed. Coordinate space is also explicit: the default `unknown` does not mean millimetres.

## 1. Read a fixed installed fixture

The CSV is project-generated CC0 teaching data. Its licence, version, size, and SHA-256 are recorded in the installed `tutorial-fixtures-6.0.csv`.

```{r read-fixture}
fixture_path <- system.file(
  "extdata", "golden", "tiny-point-grid.csv",
  package = "neurogeo",
  mustWork = TRUE
)
fixture <- read.csv(fixture_path)
fixture
```

Do not infer row order from similar coordinates. `element_id` is the stable source identifier here; row *i* of `values` must always belong to element *i* of the base.

## 2. Construct the object and declare measurement semantics

```{r construct}
point_data <- ngeo_point(
  coordinates = as.matrix(fixture[c("x_mm", "y_mm")]),
  values = cbind(signal = fixture$signal),
  measures = ngeo_measure(
    measure_id = "measure_signal",
    name = "teaching signal",
    value_type = "continuous",
    support_behavior = "intensive",
    unit = "a.u."
  ),
  coordinate_space = ngeo_coordinate_space(
    space_id = "synthetic-grid",
    kind = "unknown",
    unit = "mm"
  )
)
point_data
```

Inspect it through public accessors. A normal workflow should not mutate `$measures` or another internal table.

```{r inspect}
ngeo_base_type(point_data)
head(ngeo_base_elements(point_data))
ngeo_layers(point_data)
ngeo_measures(point_data)
ngeo_validate(point_data, "strict")
```

You should have 9 spatial elements and one layer referencing `measure_signal`.

## 3. Declare the spatial relation

Points have no intrinsic adjacency. Here Euclidean distance connects horizontal or vertical neighbors no more than 1.01 mm apart. This is an analysis choice, not a hidden property of the input.

```{r weights}
spatial_weights <- ngeo_spatial_weights(
  point_data,
  method = "distance_band",
  threshold = 1.01,
  distance_method = "euclidean",
  style = "W"
)
spatial_weights
```

KNN, another threshold, or surface geodesic distance would define a different estimand and must be reported.

## 4. Run global Moran's I

```{r moran}
global_result <- ngeo_moran(
  point_data,
  spatial_weights,
  layer = "signal",
  permutations = 199,
  seed = 2026
)
global_result
```

A positive Moran's I means neighboring signal values tend to be similar. The permutation p-value asks how often an equally extreme statistic appears when the spatial weights stay fixed and value labels are exchangeable under the stated rule. It is not the probability that a biological mechanism exists.

Check the common interpretation contract:

```{r inference-contract}
contract <- ngeo_inference_contract(global_result)
contract
```

Before reporting, verify its estimand, sampling unit, null model, metric, support, and uncertainty target. If they do not match the study design, change the analysis—not only the prose.

## 5. Inspect local patterns and history

```{r local}
local_result <- ngeo_local_moran(
  point_data,
  spatial_weights,
  layer = "signal",
  permutations = 199,
  seed = 2026,
  adjust = "BH",
  null_model = "conditional"
)
local_result[c("element_id", "local_i", "p.adjusted", "cluster")]
```

`cluster` is a Moran-quadrant description; call a local pattern significant only with the adjusted permutation evidence. The object history records construction:

```{r history}
construction <- ngeo_history(point_data)$operations[[1L]]
construction$timestamp_utc <- NULL
construction
```

## Common mistakes

- **Treating the default space as millimetres:** `ngeo_coordinate_space()` defaults to `unknown`. Confirm units from a reliable header or user declaration before physical-distance analysis.
- **Reordering values by similar names:** the package does not guess alignment. Preserve and check stable element IDs.
- **Mutating the measure table:** use `ngeo_measure()` at construction and `ngeo_update_measure()` later.
- **Treating a layer name as a permanent ID:** display names may repeat; store `layer_id` in reproducible scripts.
- **Using an experimental null as stable inference:** surface spin and the Moran eigen-sign surrogate remain uncalibrated opt-in methods and do not replace this stable permutation path.

## Reusable reporting sentence

> We analysed an intensive signal on nine ordered spatial elements using row-standardized spatial weights from a 1.01 mm Euclidean distance band. Uncertainty for global Moran's I used 199 value-label permutations with a fixed seed. Interpretation is limited to the declared base, metric, support, and null.

Next, choose the [NIfTI, surface, CIFTI, or ROI/cohort workflow](/en/tutorials/format-workflows), or continue to [neighbors, distances, and spatial weights](/en/tutorials/neighbors-and-weights).
