---
title: "Interoperability and auditable exchange"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interoperability and auditable exchange}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

NGCS 2.9 makes supported file contracts explicit. It does not add
registration, resampling, preprocessing, or dataset orchestration.

Canonical BIDS derivative names are built from validated, ordered entities.

```{r}
entities <- list(sub = "01", space = "fsLR", desc = "effect")
name <- ngeo_bids_build_name(
  entities,
  suffix = "dscalar",
  extension = ".dscalar.nii"
)
name
ngeo_bids_parse_name(name)
```

Support maps can be exchanged as checksum-verified chunks. The bundle uses
JSON plus Matrix Market and preserves the complete logical map hash.

```{r eval=FALSE}
bundle <- write_ngeo_support_bundle(
  support_map,
  "support-map.ngcs",
  chunk_size = 100000L
)
ngeo_validate_support_bundle(bundle$path)
restored <- read_ngeo_support_bundle(bundle$path)
```

CIFTI writing validates datatype and axis semantics before any final file is
published.

```{r eval=FALSE}
ngeo_validate_cifti_contract(x, "dscalar", datatype = "float64")
write_ngeo_cifti(
  x,
  "sub-01_desc-effect.dscalar.nii",
  type = "dscalar",
  datatype = "float64",
  named_map_metadata = list(list(Description = "effect estimate"))
)
```

The conformance corpus and readiness inventories are machine-readable.

```{r}
manifest <- ngeo_conformance_manifest()
manifest$corpus_version
ngeo_compatibility_matrix()
head(ngeo_api_inventory())
```
