---
title: "Interoperability and auditable exchange"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interoperability and auditable exchange}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Interoperability requires agreement on both representation and meaning. A
filename can identify a standard format, but it does not by itself guarantee
element order, spatial compatibility, measurement semantics, or file
integrity. Auditable exchange therefore combines standard encodings with
validated metadata and cryptographic hashes.

The supported exchange contracts do not add registration, resampling,
preprocessing, or dataset orchestration.

## BIDS derivative names

Canonical BIDS derivative names are built from validated, ordered entities.

```{r}
entities <- list(sub = "01", space = "fsLR", desc = "effect")
name <- ngeo_bids_build_name(
  entities,
  suffix = "dscalar",
  extension = ".dscalar.nii"
)
name
ngeo_bids_parse_name(name)
```

## Sparse support-map exchange

Support layers can be exchanged as checksum-verified chunks. The bundle uses
JSON plus Matrix Market and preserves the complete logical map hash.

```{r eval=FALSE}
bundle <- write_ngeo_support_bundle(
  support_map,
  "support-map.ngcs",
  chunk_size = 100000L
)
ngeo_validate_support_bundle(bundle$path)
restored <- read_ngeo_support_bundle(bundle$path)
```

## CIFTI output contracts

CIFTI writing validates datatype and axis semantics before any final file is
published.

```{r eval=FALSE}
ngeo_validate_cifti_contract(x, "dscalar", datatype = "float64")
write_ngeo_cifti(
  x,
  "sub-01_desc-effect.dscalar.nii",
  type = "dscalar",
  datatype = "float64",
  named_map_metadata = list(list(Description = "effect estimate"))
)
```

## Scope

Installed conformance fixtures remain part of the package test resources.
They are an implementation verification asset, not a runtime object model.
Applications should use the format-specific validators above and
`ngeo_validate()` for NGCS objects.
