---
title: "Bounded iterative spatial models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bounded iterative spatial models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Background

Exact matrix factorization becomes expensive for large spatial domains.
Iterative linear solvers and stochastic log-determinant approximations reduce
memory and computation, but introduce convergence tolerance, truncation, and
Monte Carlo error. These numerical choices are part of the scientific result
and must be recorded.

Exact small-base modelling remains a calibration reference. Iterative or
approximate execution requires an explicit control object.

## Solver controls

```{r}
library(neurogeo)

coordinates <- as.matrix(expand.grid(x = 0:4, y = 0:4))
values <- cbind(
  response = 2 + coordinates[, 1] - 0.5 * coordinates[, 2],
  x = coordinates[, 1],
  y = coordinates[, 2]
)
data <- ngeo_point(coordinates, values = values)
spatial_weights <- ngeo_spatial_weights(
  data,
  method = "distance_band",
  threshold = 1.01,
  style = "W"
)
control <- ngeo_solver_control(
  tolerance = 1e-7,
  max_iterations = 1000,
  trace_order = 20,
  trace_probes = 60,
  seed = 3401,
  workers = 1,
  exact_threshold = 100
)
```

## Calibrate on a small problem

For a small reference, request the exact determinant while retaining
iterative SAR fitted-value solving.

```{r}
sar <- ngeo_spatial_regression_iterative(
  data,
  response = "response",
  predictors = c("x", "y"),
  spatial_weights = spatial_weights,
  model = "sar",
  control = control,
  logdet = "exact"
)
sar
sar$solve$relative_residual
```

## Approximate execution

For a larger workflow, exact/approximate selection is explicit. Approximate
results include Monte Carlo and series-truncation diagnostics.

```{r}
approximate <- ngeo_logdet_approx(
  spatial_weights,
  parameter = 0.2,
  control = ngeo_solver_control(
    trace_order = 30,
    trace_probes = 100,
    exact_threshold = 10,
    seed = 3402
  ),
  exact = FALSE
)
approximate
```

## CAR, GWR, and kriging

Iterative CAR requires symmetric spatial_weights and a declared precision.

```{r}
binary <- ngeo_spatial_weights(
  data,
  method = "distance_band",
  threshold = 1.01,
  style = "B"
)
car <- ngeo_car_iterative(
  data,
  "response",
  binary,
  type = "proper",
  rho = 0.5,
  precision = 2,
  control = control
)
car
```

GWR and kriging evaluate requested targets directly; batching is an internal
execution detail rather than a second public model API.

```{r}
gwr <- ngeo_gwr(
  data,
  "response",
  c("x", "y"),
  bandwidth = 3,
  targets = c(9, 2, 17, 5, 21)
)
gwr[, c("element_id", "fitted")]
```
