---
title: "Uncertainty-aware spatial models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Uncertainty-aware spatial models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Prediction uncertainty can combine several distinct components: measurement
error in observed values, residual process variation, uncertainty in fitted
covariance parameters, and sensitivity to support or tuning choices.
Combining them without labels makes a total variance difficult to interpret.

Every uncertainty object is therefore bound to one exact ordered base so
that covariance rows and columns cannot drift away from the observations they
describe.

## Measurement covariance

```{r}
coordinates <- as.matrix(expand.grid(x = 0:3, y = 0:3))
predictor <- coordinates[, 1] + coordinates[, 2] / 4
x <- ngeo_point(
  coordinates,
  values = cbind(
    response = 1 + 2 * predictor + sin(seq_along(predictor)) / 20,
    predictor = predictor
  ),
  measures = rbind(
    ngeo_measure(support_behavior = "intensive"),
    ngeo_measure(support_behavior = "intensive")
  )
)
measurement <- ngeo_support_covariance(
  x,
  variance = rep(0.01, nrow(coordinates))
)
```

## Variogram uncertainty

Variogram uncertainty separates the raw empirical semivariance from the
measurement-error correction and retains its parameter simulations.

```{r}
variogram <- ngeo_variogram_uncertainty(
  x,
  "response",
  measurement,
  breaks = 5,
  model = "exponential",
  nsim = 20,
  seed = 27
)
variogram$empirical
variogram$parameter_interval
```

## Kriging variance components

Kriging reports independent variance components explicitly.

```{r}
prediction <- ngeo_kriging_uncertainty(
  x,
  "response",
  variogram$fit,
  targets = 1:4,
  neighbors = 8,
  value_covariance = measurement,
  variogram_uncertainty = variogram
)
prediction[
  , c(
    "process_variance", "measurement_variance",
    "parameter_variance", "support_variance", "total_variance"
  )
]
```

## Interpretation

For GWR, alternative-bandwidth minima and maxima are sensitivity ranges, not
confidence intervals. `ngeo_spatial_regression_uncertainty()` and
`ngeo_car_uncertainty()` require the same base-bound covariance.
`ngeo_support_model_ensemble()` applies the law of total variance only to the
declared support family.
