---
title: "Subject-level multilayer inference in 100 HC and 100 SCZ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Subject-level multilayer inference in 100 HC and 100 SCZ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
dk <- ngeo_tutorial_dk_case_control(n_per_group=100)
multi <- ngeo_tutorial_dk_multilayer(dk)
x <- multi$object
```

## Scientific question and data contract

For every subject, the same DK68 base carries cortical thickness and a
simulated myelin proxy. We ask whether their within-subject spatial coupling
differs between SCZ and HC after adjusting age, sex, and site. One subject is
one independent sampling unit; 68 parcels and two modalities are repeated
spatial measurements, not replication.

```{r maps, fig.width=9, fig.height=4.5, fig.alt="DK brain maps of simulated SCZ minus HC thickness and myelin proxy", fig.cap="Both modalities are inspected on their shared cortical support before coupling inference."}
thickness_values <- setNames(
  ngeo_values(dk$difference)[, 1L],
  ngeo_base_elements(dk$difference)$region_id
)
thickness_limit <- max(abs(thickness_values), na.rm=TRUE)
thickness_map <- ngeo_cortical_map(
  dk$surface,
  values=thickness_values,
  chart="flat",
  atlas=dk$atlas,
  underlay=dk$underlay,
  underlay_palette="Grays",
  overlay_alpha=0.82,
  palette="Blue-Red 3",
  limits=c(-thickness_limit, thickness_limit),
  na_color=NA_character_,
  atlas_coverage="auto"
)
plot(
  thickness_map,
  main="Thickness: SCZ minus HC (mm)",
  boundary_color=grDevices::adjustcolor("white", 0.55),
  boundary_lwd=0.25,
  outline_lwd=1.1
)

myelin_values <- setNames(
  ngeo_values(multi$myelin_difference)[, 1L],
  ngeo_base_elements(multi$myelin_difference)$region_id
)
myelin_limit <- max(abs(myelin_values), na.rm=TRUE)
myelin_map <- ngeo_cortical_map(
  dk$surface,
  values=myelin_values,
  chart="flat",
  atlas=dk$atlas,
  underlay=dk$underlay,
  underlay_palette="Grays",
  overlay_alpha=0.82,
  palette="Blue-Red 3",
  limits=c(-myelin_limit, myelin_limit),
  na_color=NA_character_,
  atlas_coverage="auto"
)
plot(
  myelin_map,
  main="Myelin proxy: SCZ minus HC (a.u.)",
  boundary_color=grDevices::adjustcolor("white", 0.55),
  boundary_lwd=0.25,
  outline_lwd=1.1
)
```

```{r data-contract}
dim(ngeo_values(x))
table(ngeo_layers(x)$feature, ngeo_layers(x)$group)
index <- ngeo_validate_layers(
  x, required_layers=c("thickness", "myelin"), complete="error"
)
index
```

There are 68 rows and 400 layers: two aligned maps for each of 200 subjects.
The layer index verifies that neither a subject nor a modality is silently
missing.

## Fix the spatial estimand before testing groups

```{r features}
weights <- ngeo_spatial_weights(
  x, method="region_contiguity", style="B"
)
basis <- ngeo_spatial_basis(
  x, weights, support="identity", n_modes=60
)
features <- ngeo_layer_coupling(
  x, index, basis=basis,
  bands=list(low_rank=1:8, mid_rank=9:20, high_rank=21:33),
  estimands=c("same_location", "spectral_coupling", "band_energy")
)
dim(features$values)
features$values[1:6, 1:6, drop=FALSE]
```

The graph basis depends only on the DK support and graph, not diagnosis or
values. Rank bands are retained graph modes; they are not automatically
millimetre wavelengths. `spectral_coupling` and marginal `band_energy` remain
separate endpoints so an energy change is not mislabeled as coupling.

## Exchange whole subjects

```{r inference}
design <- data.frame(
  unit_id=dk$design$subject_id,
  group=dk$design$group,
  age=dk$design$age,
  sex=dk$design$sex,
  site=dk$design$site
)
schedule <- ngeo_exchangeability(
  design$unit_id, scheme="within_block", blocks=design$site,
  permutations=199, seed=5001, unit_kind="subject"
)
result <- ngeo_group_test(
  features, design, model=~ age + sex + site + group, test="group",
  exchangeability=schedule, transform="auto", adjustment="maxT"
)
result$tests[, c(
  "estimand", "band", "coefficient", "partial_r2", "p_raw", "p_maxT"
)]
result$omnibus
```

One schedule moves complete subject records within site and is reused for all
bands and endpoints. It never permutes parcels. `p_maxT` controls the declared
endpoint family; the omnibus test asks whether any endpoint in that family
differs by group under the same schedule.

## Extend to a support family

Run the feature construction independently on registered DK68,
Schaefer100/200/300, or another prespecified atlas, then pass the named feature
list to one `ngeo_group_test()` call. A common subject schedule and full-family
max-T adjustment are essential. Atlas dispersion describes sensitivity; it
does not prove parcellation invariance.

The simulator's Schaefer-sized boundaries are for operator teaching only.
Scientific cross-atlas claims require the published labels registered to the
same subjects and documented physical scale.
