---
title: "Multilayer spatial inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multilayer spatial inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Scientific question and data contract

This workflow asks whether two aligned cortical layers are coupled, at which
retained spatial scales that coupling occurs, and whether coupling differs
between subject groups. Every map belongs to one independent subject and one
layer. Vertices or regions are spatial elements, not replicated subjects.

```{r data}
library(neurogeo)

set.seed(50)
n_element <- 10L
n_subject <- 16L
unit_id <- sprintf("sub-%02d", seq_len(n_subject))
group <- factor(rep(c("control", "case"), each = n_subject / 2L))
age <- seq(20, 55, length.out = n_subject)
feature <- rep(c("thickness", "myelin"), n_subject)
subject <- rep(unit_id, each = 2L)
base <- sin(seq_len(n_element) / 2)
values <- matrix(NA_real_, n_element, 2L * n_subject)
for (i in seq_len(n_subject)) {
  values[, 2L * i - 1L] <- base + rnorm(n_element, sd = 0.12)
  strength <- if (group[[i]] == "case") 0.8 else 0.4
  values[, 2L * i] <- strength * base + rnorm(n_element, sd = 0.12)
}
maps <- data.frame(
  map_id = sprintf("map-%03d", seq_len(ncol(values))),
  name = paste(subject, feature, sep = "_"),
  subject_id = subject, feature = feature
)
measure <- ngeo_measure(spatial_semantics = "intensive", units = "a.u.")
x <- ngeo_points(
  cbind(x = seq_len(n_element), y = 0), values = values, maps = maps,
  measures = measure[rep.int(1L, ncol(values)), , drop = FALSE]
)
index <- ngeo_validate_layers(
  x, required_layers = c("thickness", "myelin"), complete = "error"
)
```

## Fixed spatial estimand

The graph basis below depends on coordinates, topology, and declared support,
not on the values that will later be tested. The two named bands are
rank-matched retained modes; they are not automatically physical wavelengths.

```{r features}
weights <- ngeo_weights(
  x, method = "distance_band", threshold = 1.01, style = "B"
)
basis <- ngeo_spatial_basis(
  x, weights, support = "identity", n_modes = n_element - 1L
)
features <- ngeo_layer_coupling(
  x, index, basis = basis,
  bands = list(low_rank = 1:4, high_rank = 5:9),
  estimands = c("same_location", "spectral_coupling", "band_energy")
)
features
```

`spectral_coupling` is normalized cross-layer covariance in a band.
`band_energy_x` and `band_energy_y` are marginal layer structure. Keeping them
as separate endpoints prevents an energy change from being described as a
coupling change.

## Subject-level inference

The exchangeability object transforms whole subjects. One schedule is reused
for every endpoint; no vertex-wise pseudo-replication occurs.

```{r inference}
design <- data.frame(unit_id = unit_id, group = group, age = age)
schedule <- ngeo_exchangeability(
  unit_id, scheme = "free", permutations = 99L, seed = 5001L
)
result <- ngeo_group_test(
  features, design, model = ~ age + group, test = "group",
  exchangeability = schedule, transform = "auto", adjustment = "maxT"
)
result$tests[, c(
  "estimand", "band", "coefficient", "partial_r2", "p_raw", "p_maxT"
)]
result$omnibus
```

The valid inference unit is the independent subject. The result supports a
claim about the declared population only when residual exchangeability and the
subject design are defensible. It does not establish causality, registration
quality, equal physical scale across atlases, or parcellation invariance.

## Declared support family

Construct features independently on every atlas or support, then pass the
named list to the same group call. This produces one common schedule and one
full-family max-T calculation. Support dispersion remains descriptive.

```{r support-family, eval=FALSE}
family_result <- ngeo_group_test(
  features = list(DK68 = dk_features, Schaefer100 = schaefer_features),
  data = design, model = ~ age + group, test = "group",
  exchangeability = schedule
)
family_result$support$stability
```

Use `ngeo_qc()` on the inputs and inspect basis diagnostics, missing endpoints,
schedule hashes, and support scale labels before interpreting the tests.
