---
title: "Cortical neighbors and weights on a DK68 case-control map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cortical neighbors and weights on a DK68 case-control map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
dk <- ngeo_tutorial_dk_case_control()
x <- dk$difference
```

## The map whose neighborhood we are defining

`x` is the SCZ-minus-HC cortical-thickness contrast on DK68. A neighborhood is
part of the estimand: cortical contiguity, centroid KNN, and surface geodesic
distance answer different questions.

```{r input-map, fig.width=9, fig.height=4.6, fig.alt="SCZ minus HC cortical thickness on DK68", fig.cap="Always orient the spatial relation on the brain map it will analyze."}
contrast_values <- setNames(
  ngeo_values(x)[, "SCZ_minus_HC"],
  ngeo_base_elements(x)$region_id
)
contrast_limit <- max(abs(contrast_values), na.rm=TRUE)
contrast_map <- ngeo_cortical_map(
  dk$surface,
  values=contrast_values,
  chart="flat",
  atlas=dk$atlas,
  underlay=dk$underlay,
  underlay_palette="Grays",
  overlay_alpha=0.82,
  palette="Blue-Red 3",
  limits=c(-contrast_limit, contrast_limit),
  na_color=NA_character_,
  atlas_coverage="auto"
)
plot(
  contrast_map,
  main="SCZ minus HC cortical thickness (mm)",
  boundary_color=grDevices::adjustcolor("white", 0.55),
  boundary_lwd=0.25,
  outline_lwd=1.1
)
```

## Use the declared DK graph

The teaching object stores a symmetric local graph derived from the atlas
drawing. It keeps hemispheres disconnected. A real surface analysis should
derive topology from registered mesh faces or a documented geodesic rule.

```{r weights}
binary <- ngeo_spatial_weights(
  x, method="region_contiguity", style="B"
)
row_standardized <- ngeo_spatial_weights(
  x, method="region_contiguity", style="W"
)
c(
  parcels = nrow(ngeo_values(x)),
  undirected_edges = sum(binary$raw_matrix != 0) / 2,
  isolated_parcels = sum(Matrix::rowSums(binary$raw_matrix != 0) == 0)
)
summary(Matrix::rowSums(binary$raw_matrix != 0))
```

Binary weights count every neighboring parcel equally. Row-standardized
weights give every parcel total outgoing weight one, so a parcel with many
neighbors contributes less per edge. This normalization changes the statistic
and must be reported.

```{r weight-diagnostic, fig.width=7, fig.height=4, fig.alt="Histogram of DK parcel neighbor counts", fig.cap="A degree diagnostic accompanies the brain map; it does not replace it."}
plot(binary)
```

## Global and local spatial association

```{r inference}
global <- ngeo_moran(
  x, row_standardized, "SCZ_minus_HC",
  permutations=499, seed=2026
)
local <- ngeo_local_moran(
  x, row_standardized, "SCZ_minus_HC",
  permutations=499, seed=2026, adjust="BH",
  null_model="conditional"
)
global
```

Put local Moran's I back on the same cortical support instead of reading it as
an unlocated vector.

```{r local-map, fig.width=9, fig.height=4.6, fig.alt="Local Moran I for the case-control contrast plotted on DK68", fig.cap="Local statistics are returned to their cortical parcels for interpretation."}
local_map <- ngeo_parcellation(
  dk$regions, values=cbind(local_I=local$local_i),
  centroid=dk$centroids,
  support_size=dk$support_size, adjacency=dk$adjacency,
  measures=ngeo_measure(name="local Moran I", support_behavior="intensive",
                        unit="dimensionless"),
  coordinate_space=ngeo_spatial_base(dk$cohort)$coordinate_space
)
local_values <- setNames(
  ngeo_values(local_map)[, "local_I"],
  ngeo_base_elements(local_map)$region_id
)
local_limit <- max(abs(local_values), na.rm=TRUE)
local_brain_map <- ngeo_cortical_map(
  dk$surface,
  values=local_values,
  chart="flat",
  atlas=dk$atlas,
  underlay=dk$underlay,
  underlay_palette="Grays",
  overlay_alpha=0.82,
  palette="Blue-Red 3",
  limits=c(-local_limit, local_limit),
  na_color=NA_character_,
  atlas_coverage="auto"
)
plot(
  local_brain_map,
  main="Local Moran's I of SCZ minus HC",
  boundary_color=grDevices::adjustcolor("white", 0.55),
  boundary_lwd=0.25,
  outline_lwd=1.1
)
```

The permutation null randomizes values over parcels while holding the graph
fixed. It characterizes spatial structure of the contrast and is not a
subject-level diagnosis test. For that, exchange whole subjects under the
declared case-control design.

## Reportable specification

Record the base (DK68), graph source, metric, neighborhood rule, normalization,
island handling, diagonal convention, permutation count, seed, and independent
sampling unit. Changing any of these can change the scientific result.
