---
title: "Neighbors and weights"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Neighbors and weights}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.retina = 2,
  out.width = "100%"
)
library(neurogeo)
```

## Background

Spatial statistics require a declared answer to “which locations can
influence one another?” A weights matrix \(W\) encodes that answer: row \(i\)
contains the influence assigned to neighbors of element \(i\). Different
construction rules imply different scientific neighborhoods and can change
the result even when the observed values are unchanged.

Topology-derived relationships are sparse. Surface adjacency follows mesh
edges, volume adjacency uses active-mask-limited 6/18/26 connectivity, and
grayordinate adjacency remains block diagonal by default.

## Construct and inspect weights

```{r}
coordinates <- as.matrix(expand.grid(x = 0:2, y = 0:2))
x <- ngeo_points(
  coordinates,
  values = cbind(signal = c(1, 2, 3, 2, 4, 7, 3, 7, 9)),
  measures = ngeo_measure(spatial_semantics = "intensive")
)
w <- ngeo_weights(
  x,
  method = "distance_band",
  threshold = 1.01,
  style = "W"
)
w
w$diagnostics
```

Inspect the graph before using it in a statistic. The node fill shows the
signal while each segment shows one distance-band connection.

```{r distance-network, fig.alt="A three by three spatial graph with horizontal and vertical neighbor connections", fig.cap="The distance-band graph. A visual check catches missing components and unintended diagonal links more quickly than a matrix printout."}
W <- as.matrix(w$matrix)
edges <- which(
  upper.tri(W) & ((W > 0) | (t(W) > 0)),
  arr.ind = TRUE
)
node_colors <- grDevices::hcl.colors(9, "YlOrRd", rev = TRUE)[
  cut(x$values[, "signal"], breaks = 9, labels = FALSE, include.lowest = TRUE)
]
plot(coordinates, type = "n", asp = 1, xlab = "x", ylab = "y",
     main = "Distance-band neighbors")
segments(
  coordinates[edges[, 1], 1], coordinates[edges[, 1], 2],
  coordinates[edges[, 2], 1], coordinates[edges[, 2], 2],
  col = "#8aa6a3", lwd = 2
)
points(coordinates, pch = 21, bg = node_colors, col = "white", cex = 2.8)
text(coordinates, labels = seq_len(nrow(coordinates)), font = 2)
```

## Weight styles and symmetry

`style = "W"` row-standardizes the relation, `style = "B"` makes it binary,
and `style = "none"` retains raw edge or kernel weights. KNN construction
requires an explicit symmetry policy.

```{r}
knn <- ngeo_weights(
  x,
  method = "knn",
  k = 2,
  symmetry = "union",
  style = "B"
)
knn
```

The degree distributions make the consequences of the two construction
rules visible. Boundary nodes have fewer distance-band neighbors, whereas KNN
starts from a requested neighbor count and then applies the symmetry policy.

```{r compare-degrees, fig.width=8, fig.height=4.5, fig.alt="Side-by-side bar charts comparing node degree under distance-band and K nearest-neighbor weights", fig.cap="Neighbor count is a model choice, not a harmless implementation detail."}
old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
barplot(
  Matrix::rowSums(w$raw_matrix != 0),
  names.arg = seq_len(nrow(coordinates)),
  col = "#26a69a",
  border = NA,
  xlab = "Element",
  ylab = "Neighbor count",
  main = "Distance band"
)
barplot(
  Matrix::rowSums(knn$raw_matrix != 0),
  names.arg = seq_len(nrow(coordinates)),
  col = "#e29b45",
  border = NA,
  xlab = "Element",
  ylab = "Neighbor count",
  main = "KNN union"
)
par(old_par)
```

## Distance and spatial association

Distances require explicit sources and targets; the API guards against
accidental dense all-pairs allocation.

```{r}
ngeo_distance(x, from = 1, to = c(2, 5, 9), metric = "euclidean")
```

The same weights object can drive global and local statistics.

```{r}
moran <- ngeo_moran(x, w, "signal")
lisa <- ngeo_local_moran(x, w, "signal")
moran
head(lisa)
```

```{r association-plots, fig.width=8, fig.height=4.5, fig.alt="Side-by-side Moran and local Moran scatterplots", fig.cap="Global and local association diagnostics use the same declared weights object."}
old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
plot(moran, pch = 21, bg = "#26a69a", col = "white", cex = 1.4)
plot(lisa, pch = 21, bg = "#e29b45", col = "white", cex = 1.4)
par(old_par)
```
