---
title: "Neighbors and weights"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Neighbors and weights}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

Topology-derived relationships are sparse. Surface adjacency follows mesh
edges, volume adjacency uses active-mask-limited 6/18/26 connectivity, and
grayordinate adjacency remains block diagonal by default.

```{r}
coordinates <- as.matrix(expand.grid(x = 0:2, y = 0:2))
x <- ngeo_points(
  coordinates,
  values = cbind(signal = c(1, 2, 3, 2, 4, 7, 3, 7, 9)),
  measures = ngeo_measure(spatial_semantics = "intensive")
)
w <- ngeo_weights(
  x,
  method = "distance_band",
  threshold = 1.01,
  style = "W"
)
w
w$diagnostics
```

`style = "W"` row-standardizes the relation, `style = "B"` makes it binary,
and `style = "none"` retains raw edge or kernel weights. KNN construction
requires an explicit symmetry policy.

```{r}
knn <- ngeo_weights(
  x,
  method = "knn",
  k = 2,
  symmetry = "union",
  style = "B"
)
knn
```

Distances require explicit sources and targets; the API guards against
accidental dense all-pairs allocation.

```{r}
ngeo_distance(x, from = 1, to = c(2, 5, 9), metric = "euclidean")
```

The same weights object can drive global and local statistics.

```{r}
ngeo_moran(x, w, "signal")
head(ngeo_local_moran(x, w, "signal"))
```
