---
title: "Parcellation and aggregation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parcellation and aggregation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.retina = 2,
  out.width = "100%"
)
library(neurogeo)
```

## Background

A parcellation assigns fine-scale elements to regions. Aggregation then
changes the observational support from vertices, voxels, or points to those
regions. The correct reduction depends on what a variable measures: averaging
a thickness is meaningful, while averaging a regional count would destroy
its total.

Aggregation follows declared measurement semantics. Intensive quantities
use support-weighted means, extensive quantities and counts use sums,
categorical values use a declared mode rule, and unknown semantics require
an explicit function.

## Construct a partition

```{r}
coordinates <- matrix(
  c(
    0, 0, 0,
    1, 0, 0,
    1, 1, 0,
    0, 1, 0
  ),
  ncol = 3,
  byrow = TRUE
)
faces <- matrix(c(1, 2, 3, 1, 3, 4), ncol = 3, byrow = TRUE)
values <- cbind(
  thickness = c(1, 2, 3, 4),
  surface_area = c(1 / 3, 1 / 6, 1 / 3, 1 / 6)
)
surface <- ngeo_surface(
  coordinates,
  faces,
  values = values,
  measures = rbind(
    ngeo_measure(spatial_semantics = "intensive", units = "mm"),
    ngeo_measure(spatial_semantics = "extensive", units = "mm2")
  )
)
partition <- ngeo_partition(surface, c("A", "A", "B", "B"))
regional <- ngeo_aggregate(surface, partition)
regional$values
regional$domain$support_size
```

Plot the surface map and the partition sizes before interpreting the regional
values. The first view checks spatial alignment; the second checks whether one
parcel unexpectedly dominates the aggregation.

```{r parcellation-plots, fig.width=8, fig.height=4.5, fig.alt="Surface thickness values and a bar chart of partition region sizes", fig.cap="Geometry and partition diagnostics should precede aggregation."}
old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
plot(surface, map = "thickness", show_edges = TRUE, palette = "Viridis")
plot(partition, col = c("#3b82a0", "#e29b45"), border = NA)
par(old_par)
```

## Check measurement semantics

The extensive map is conserved:

```{r}
sum(surface$values[, "surface_area"])
sum(regional$values[, "surface_area"])
```

```{r semantics-comparison, fig.width=8, fig.height=4.5, fig.alt="Two grouped bar charts comparing source elements with aggregated regional thickness and surface area", fig.cap="Intensive thickness is averaged, whereas extensive surface area is conserved and summed."}
old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
barplot(
  rbind(
    source_mean = tapply(surface$values[, "thickness"],
                         partition$membership, mean),
    regional = regional$values[, "thickness"]
  ),
  beside = TRUE,
  col = c("#8ecdc4", "#176b63"),
  border = NA,
  ylab = "Thickness (mm)",
  main = "Intensive"
)
legend("topleft", legend = c("Source mean", "Regional"),
       fill = c("#8ecdc4", "#176b63"), bty = "n", cex = 0.8)
barplot(
  rbind(
    source_sum = tapply(surface$values[, "surface_area"],
                        partition$membership, sum),
    regional = regional$values[, "surface_area"]
  ),
  beside = TRUE,
  col = c("#f3c985", "#d97b29"),
  border = NA,
  ylab = expression(mm^2),
  main = "Extensive"
)
legend("topleft", legend = c("Source sum", "Regional"),
       fill = c("#f3c985", "#d97b29"), bty = "n", cex = 0.8)
par(old_par)
```

## Labels, boundaries, and identity

Label inputs can be used directly by name. Background and medial-wall values
are never guessed; declare them explicitly.

```{r labels, eval=FALSE}
partition <- ngeo_partition(surface, "aparc", background = 0)
boundary <- ngeo_boundary(surface, partition)
region_adjacency <- ngeo_region_adjacency(surface, partition)
```

Partitions carry the base-domain hash, so they cannot be silently reused
after a geometry or indexing change.
