---
title: "Parcellation and aggregation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parcellation and aggregation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

Aggregation follows declared measurement semantics. Intensive quantities
use support-weighted means, extensive quantities and counts use sums,
categorical values use a declared mode rule, and unknown semantics require
an explicit function.

```{r}
coordinates <- matrix(
  c(
    0, 0, 0,
    1, 0, 0,
    1, 1, 0,
    0, 1, 0
  ),
  ncol = 3,
  byrow = TRUE
)
faces <- matrix(c(1, 2, 3, 1, 3, 4), ncol = 3, byrow = TRUE)
values <- cbind(
  thickness = c(1, 2, 3, 4),
  surface_area = c(1 / 3, 1 / 6, 1 / 3, 1 / 6)
)
surface <- ngeo_surface(
  coordinates,
  faces,
  values = values,
  measures = rbind(
    ngeo_measure(spatial_semantics = "intensive", units = "mm"),
    ngeo_measure(spatial_semantics = "extensive", units = "mm2")
  )
)
partition <- ngeo_partition(surface, c("A", "A", "B", "B"))
regional <- ngeo_aggregate(surface, partition)
regional$values
regional$domain$support_size
```

The extensive map is conserved:

```{r}
sum(surface$values[, "surface_area"])
sum(regional$values[, "surface_area"])
```

Label inputs can be used directly by name. Background and medial-wall values
are never guessed; declare them explicitly.

```{r labels, eval=FALSE}
partition <- ngeo_partition(surface, "aparc", background = 0)
boundary <- ngeo_boundary(surface, partition)
region_adjacency <- ngeo_region_adjacency(surface, partition)
```

Partitions carry the base-domain hash, so they cannot be silently reused
after a geometry or indexing change.
