---
title: "Unified scientific quality control"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unified scientific quality control}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Validation and QC answer different questions

`ngeo_validate()` enforces NGCS structural invariants. `ngeo_qc()` reports
conditions that can be valid but require a scientific decision. QC does not
repair data or infer missing metadata.

```{r basic}
x <- ngeo_point(
  matrix(c(0, 0, 1, 0, 2, 0), ncol = 2, byrow = TRUE),
  values = cbind(signal = c(1, 2, 3), constant = c(5, 5, 5))
)

qc <- ngeo_qc(x)
qc
qc$checks
qc$layer_summary
```

```{r plot, fig.width=6, fig.height=4}
plot(qc)
```

The status vocabulary is `pass`, `info`, `warning`, `not_evaluated`, and
`not_applicable`. A large values block is never read beyond
`max_value_cells`:

```{r budget}
ngeo_qc(x, max_value_cells = 2)$checks
```

An aligned support map adds sparse coverage and conservation checks:

```{r support}
atlas <- ngeo_atlas_map(
  x,
  c("A", "A", NA),
  source_support = rep(1, 3)
)
ngeo_qc(x, support_map = atlas)$checks
```

QC does not perform registration, resampling, preprocessing, outlier
removal, or statistical inference. A passing report is not a substitute for
the study design and measurement assumptions.
