---
title: "Quality control for DK68 and vertex case-control data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality control for DK68 and vertex case-control data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
dk <- ngeo_tutorial_dk_case_control()
vertex <- ngeo_tutorial_vertex_case_control()
```

## Validation and QC answer different questions

`ngeo_validate()` checks structural invariants: 68 DK rows align across all 200
subjects, layer metadata align with columns, and measure semantics are valid.
`ngeo_qc()` reports scientifically relevant conditions without repairing them.

```{r validate}
ngeo_validate(dk$cohort, "strict")
qc <- ngeo_qc(dk$cohort)
qc$summary
qc$checks
head(qc$layer_summary)
```

## QC remains spatially oriented

Inspect a subject map and the case-control contrast alongside the tabular QC
report. A global pass cannot reveal a localized but anatomically implausible
pattern.

```{r brain-qc, fig.width=9, fig.height=4.6, fig.alt="One simulated HC cortical thickness map on DK68 for visual QC", fig.cap="Structural validation and a brain plot answer complementary QC questions."}
ngeo_tutorial_plot_dk(
  dk$cohort, "sub-001", title="Visual QC: sub-001 cortical thickness"
)
```

```{r qc-plot, fig.width=7, fig.height=4, fig.alt="Summary plot of neurogeo QC statuses", fig.cap="The status plot summarizes machine-readable checks; it does not localize values on cortex."}
plot(qc)
```

## Bound the values scan explicitly

For a large cohort, `max_value_cells` limits how many cells QC may inspect. A
bounded scan reports what was not evaluated; it never silently treats an
unread block as passing.

```{r budget}
bounded_qc <- ngeo_qc(dk$cohort, max_value_cells=1000)
bounded_qc$checks
```

## Audit support coverage on vertex data

The complete Schaefer200-sized teaching map assigns every one of 672 vertices
to exactly one target parcel. We then remove an anterior-temporal subset to
show partial coverage on the brain and in sparse diagnostics.

```{r support}
complete_map <- ngeo_atlas_map(
  vertex$difference, vertex$supports$Schaefer200,
  source_support=ngeo_vertex_area(vertex$difference)
)
partial_labels <- vertex$supports$Schaefer200
partial_labels[vertex$truth < -0.10] <- NA_character_
partial_map <- ngeo_atlas_map(
  vertex$difference, partial_labels,
  source_support=ngeo_vertex_area(vertex$difference)
)
ngeo_qc(vertex$difference, support_map=partial_map)$checks
```

```{r coverage-brain, fig.width=9, fig.height=4.5, fig.alt="Vertex coverage mask with an anterior temporal missing region on bilateral cortex", fig.cap="The missing support is visible on the brain and quantified by the support-map QC."}
coverage_surface <- ngeo_tutorial_surface_map(
  vertex$difference, as.numeric(!is.na(partial_labels)),
  "support coverage", unit="binary"
)
plot(coverage_surface, chart="flat", show_edges=FALSE, palette="Blues 3")
```

QC does not perform registration, preprocessing, outlier removal, or
inference. A passing report is evidence about the checks that ran, not proof
that the study design or anatomy is correct.
