---
title: "Reading neuroimaging data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading neuroimaging data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

A neuroimaging file contains more than a numeric array. Interpretation also
depends on element order, coordinate transforms, brain structures, map names,
label tables, units, and the source file identity. Losing any of these can
produce a numerically valid object in the wrong space or with misaligned
values.

Readers therefore preserve source indexing, coordinate transforms, map
metadata, labels, and provenance. They do not invoke FreeSurfer, FSL, or
Connectome Workbench, and they do not perform registration or resampling.

## Format-specific readers

```{r readers, eval=FALSE}
volume <- read_ngeo_nifti("sub-01_T1w.nii.gz")

surface <- read_ngeo_gifti(
  geometry = "sub-01.L.midthickness.surf.gii",
  data = c(thickness = "sub-01.L.thickness.shape.gii"),
  labels = c(aparc = "sub-01.L.aparc.label.gii")
)

dense <- read_ngeo_cifti(
  "sub-01_task-rest.dtseries.nii",
  surfaces = list(
    CORTEX_LEFT = "sub-01.L.midthickness.surf.gii",
    CORTEX_RIGHT = "sub-01.R.midthickness.surf.gii"
  )
)

freesurfer <- read_ngeo_freesurfer(
  geometry = "lh.white",
  data = c(thickness = "lh.thickness"),
  labels = "lh.aparc.annot"
)
```

## Dispatch and validation

`read_ngeo()` dispatches by standard filename suffix. For ambiguous MGH/MGZ
inputs, use the format-specific reader and state the intended domain.

```{r dispatch, eval=FALSE}
x <- read_ngeo("sub-01_space-MNI152_desc-stat.nii.gz")
x <- read_ngeo_freesurfer("lh.metric.mgh", domain = "surface",
                          surface = "lh.white")
```

After import, validate before analysis and declare any measurement semantics
that the source did not encode.

```{r workflow, eval=FALSE}
ngeo_validate(surface, "strict")
surface$measures$spatial_semantics <- "intensive"
surface$measures$units <- "mm"
w <- ngeo_weights(surface, method = "mesh_contiguity", style = "W")
ngeo_moran(surface, w, "thickness", permutations = 999, seed = 2026)
```

## Scope

BIDS sidecars and relevant space entities are parsed when present, but
dataset orchestration remains outside the package. Import establishes an
auditable object contract; it does not establish anatomical correspondence
between datasets.
