---
title: "Real-world support mapping"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real-world support mapping}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

A support map is valid only when the correspondence between source and target
domains is already known. That evidence may be a registered surface, a voxel
affine, or atlas probabilities defined on the same ordered elements.
Interpolation constructs weights from that evidence; it does not discover
anatomical correspondence.

NGCS separates the evidence that two domains correspond from the sparse
operator used to change support. `neurogeo` builds operators from known
registrations, affines, and aligned atlas data; it does not estimate those
relationships.

## A known surface registration

Both surfaces below declare one registered space. In practice the named
coordinate sets can be registered spheres supplied by a standard workflow.

```{r}
coordinates <- matrix(
  c(0, 0, 0, 1, 0, 0, 1, 1, 0, 0, 1, 0),
  ncol = 3,
  byrow = TRUE
)
faces <- matrix(c(1, 2, 3, 1, 3, 4), ncol = 3, byrow = TRUE)
space <- ngeo_space(
  "declared-registration",
  kind = "surface",
  structure = "CORTEX_LEFT"
)
source <- ngeo_surface(
  coordinates,
  faces,
  values = cbind(thickness = c(2.1, 2.4, 2.2, 2.5)),
  measures = ngeo_measure(spatial_semantics = "intensive"),
  space = space
)
target <- ngeo_surface(coordinates, faces, space = space)

surface_map <- ngeo_surface_registration_map(
  source,
  target,
  method = "barycentric"
)
ngeo_support_diagnostics(surface_map)
resampled <- ngeo_change_support(source, target, surface_map)
resampled$values
```

If spaces differ or are unknown, `registration` must explicitly identify
the already-known relationship. This argument does not run registration.

## A known affine grid

Nearest and trilinear builders transform source voxel centres through the
stored affines while respecting each object's source index base.

```{r}
volume_space <- ngeo_space("template-grid", kind = "volume")
source_grid <- ngeo_volume(
  dim = c(2, 2, 2),
  affine = diag(4),
  space = volume_space,
  index_base = "zero"
)
target_grid <- ngeo_volume(
  dim = c(2, 2, 2),
  affine = diag(4),
  space = volume_space,
  index_base = "zero"
)
grid_map <- ngeo_affine_grid_map(
  source_grid,
  target_grid,
  method = "trilinear"
)
ngeo_support_diagnostics(grid_map)
```

`ngeo_voxel_overlap_map()` computes exact fractional overlap only for
axis-aligned grids. Rotation and shear are rejected rather than silently
approximated.

## Hard and probabilistic atlases

Hard labels create a crisp operator and, when needed, a bound region target.

```{r}
atlas <- ngeo_atlas_map(
  source,
  c("anterior", "anterior", "posterior", "posterior")
)
regional <- ngeo_change_support(source, atlas$target, atlas)
regional$values
```

Probabilities are supplied as source-element by region. The builder performs
the single required transpose into the normative target-by-source operator.

```{r}
probability <- matrix(
  c(1, 0, 0.7, 0.3, 0.3, 0.7, 0, 1),
  ncol = 2,
  byrow = TRUE,
  dimnames = list(NULL, c("anterior", "posterior"))
)
soft <- ngeo_probabilistic_atlas_map(source, probability)
ngeo_support_entropy(soft)
```

## Sensitivity and uncertainty

Alternative operators can be compared only when their ordered source and
target domains match.

```{r}
soft_same_target <- ngeo_probabilistic_atlas_map(
  source,
  probability,
  target = atlas$target
)
sensitivity <- ngeo_support_sensitivity(
  source,
  atlas$target,
  list(hard = atlas, soft = soft_same_target)
)
sensitivity
```

An operator with independent entry variances can be sampled under an
explicit truncation and normalization model:

```{r}
uncertain <- soft
uncertain$weight_variance <- methods::as(
  uncertain$operator * 0.0025,
  "dgCMatrix"
)
draws <- ngeo_support_monte_carlo(
  uncertain,
  nsim = 10,
  seed = 2026
)
draws
```

The independence and Gaussian assumptions are recorded. They are not a
substitute for a registration or segmentation uncertainty model.

## Sparse exchange

`write_ngeo_support_map()` writes a Matrix Market operator and JSON sidecar.
An optional aligned variance operator is a second Matrix Market file.
`read_ngeo_support_map()` verifies the stored integrity hash.
