---
title: "Auditable replay and derivative artifacts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auditable replay and derivative artifacts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Background

Reproducibility requires more than saving final numbers. A replay record must
identify exact inputs, ordered operations, parameters, intermediate hashes,
the execution environment, and expected outputs. Restricting the operation
vocabulary makes the record inspectable and prevents a manifest from becoming
an arbitrary-code container.

A small deterministic analysis can be recorded as data. This is a
verification layer, not a general workflow engine: manifests cannot contain R
functions or arbitrary code.

## Record and verify an analysis

```{r}
library(neurogeo)

x <- ngeo_point(
  cbind(x = 1:4, y = 0, z = 0),
  values = cbind(t0 = 1:4, t1 = 2 * (1:4), t2 = 3 * (1:4))
)
x <- ngeo_set_time_axis(
  x,
  ngeo_time_axis(start = 0, step = 1, n = 3, unit = "day"),
  temporal_semantics = "instantaneous"
)

steps <- list(
  ngeo_replay_step(
    "window", "ngeo_time_slice", c(x = "source"),
    list(index = 1:2)
  ),
  ngeo_replay_step(
    "trend", "ngeo_temporal_trend", c(x = "window")
  )
)
manifest <- ngeo_record_replay(
  list(source = x), steps, outputs = "trend"
)
verified <- ngeo_replay(manifest, list(source = x))
verified$verified
```

## Logical identity

The logical hash covers the ordered base, aligned values, measurement
semantics, layers, labels, and file identity. Execution timestamps do not
change it. Replay checks the environment and every input, intermediate, and
final hash before returning verified outputs.

## Publish related artifacts

Related derivatives can be published as one complete directory transition.

```{r eval=FALSE}
batch <- ngeo_write_artifact_batch(
  "derivatives/neurogeo/sub-01",
  c("sub-01_metric.tsv", "sub-01_metric.json"),
  list(
    function(path) write.table(distance_method, path, sep = "\t"),
    function(path) jsonlite::write_json(sidecar, path)
  ),
  roles = c("distance_method", "sidecar"),
  entities = list(subject = "01", datatype = "anat")
)
ngeo_read_artifact_batch("derivatives/neurogeo/sub-01")
```

Artifacts use root-relative paths and content SHA-256 values. Batch scope is
fixed to `derivative_only`; neurogeo does not discover or orchestrate an
entire BIDS dataset.
