---
title: "Reproducible replay of a DK case-control cortical selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducible replay of a DK case-control cortical selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
dk <- ngeo_tutorial_dk_case_control()
```

## Record data identity and the intended cortical subset

Reproducibility is more than saving the final numbers. A replay record identifies
the exact 68-by-200 input, ordered operations, parameters, environment, and
expected hashes. Here a preregistered cortical set contains bilateral superior
temporal, insular, anterior cingulate, and medial orbitofrontal parcels.

```{r selection}
selected <- which(dk$regions$region %in% c(
  "superior temporal", "insula", "caudal anterior cingulate",
  "rostral anterior cingulate", "medial orbitofrontal"
))
steps <- list(ngeo_replay_step(
  "cortical_set", "ngeo_subset", c(x="source"),
  list(elements=selected, layers=seq_len(200))
))
manifest <- ngeo_record_replay(
  list(source=dk$cohort), steps, outputs="cortical_set"
)
verified <- ngeo_replay(manifest, list(source=dk$cohort))
verified$verified
manifest$canonical_sha256
```

## Verify the output on the brain

The replayed object retains all 200 subject layers but only the declared
parcels. Unselected DK cortex is gray, which makes the recorded spatial scope
immediately visible.

```{r replay-brain, fig.width=9, fig.height=4.6, fig.alt="Replayed preregistered DK cortical subset for one HC subject", fig.cap="A replay hash verifies identity; the brain plot verifies that the selected spatial scope is the intended one."}
replayed <- verified$outputs$cortical_set
ngeo_tutorial_plot_dk(
  replayed, "sub-001", title="Verified replay: preregistered cortical set"
)
```

The logical hash covers ordered base elements, values, layer and measure
metadata, labels, and file identity. Timestamps do not change it. Replay checks
the input, intermediate, and output hash before returning `verified = TRUE`.
The restricted operation vocabulary prevents the manifest from becoming an
arbitrary-code container.

## Publish related derivatives

```{r publish, eval=FALSE}
batch <- ngeo_write_artifact_batch(
  "derivatives/neurogeo/group-scz-vs-hc",
  c("group_desc-SCZminusHC_thickness.tsv", "group_desc-SCZminusHC_thickness.json"),
  list(
    function(path) write.table(parcel_effects, path, sep="\t", row.names=FALSE),
    function(path) jsonlite::write_json(sidecar, path, auto_unbox=TRUE)
  ),
  roles=c("effect_map", "sidecar"),
  entities=list(analysis="SCZminusHC", datatype="anat")
)
```

Artifacts use root-relative paths and SHA-256 content hashes. `neurogeo`
publishes the declared derivative batch; it does not discover or orchestrate
an entire BIDS dataset.
