---
title: "Auditable replay and derivative artifacts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auditable replay and derivative artifacts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

NGCS 3.5 records a small, deterministic neurogeo analysis as data. It is a
verification layer, not a general workflow engine: manifests cannot contain
R functions or arbitrary code.

```{r}
library(neurogeo)

x <- ngeo_points(
  cbind(x = 1:4, y = 0, z = 0),
  values = cbind(t0 = 1:4, t1 = 2 * (1:4), t2 = 3 * (1:4))
)
x <- ngeo_set_time_axis(
  x,
  ngeo_time_axis(start = 0, step = 1, n = 3, unit = "day"),
  temporal_semantics = "instantaneous"
)

steps <- list(
  ngeo_replay_step(
    "window", "ngeo_time_slice", c(x = "source"),
    list(index = 1:2)
  ),
  ngeo_replay_step(
    "trend", "ngeo_temporal_trend", c(x = "window")
  )
)
manifest <- ngeo_record_replay(
  list(source = x), steps, outputs = "trend"
)
verified <- ngeo_replay(manifest, list(source = x))
verified$verified
```

The logical hash covers the ordered domain, aligned values, measurement
semantics, maps, labels, and file identity. Execution timestamps do not
change it. Replay checks the environment and every input, intermediate, and
final hash before returning verified outputs.

Related derivatives can be published as one complete directory transition.

```{r eval=FALSE}
batch <- ngeo_write_artifact_batch(
  "derivatives/neurogeo/sub-01",
  c("sub-01_metric.tsv", "sub-01_metric.json"),
  list(
    function(path) write.table(metric, path, sep = "\t"),
    function(path) jsonlite::write_json(sidecar, path)
  ),
  roles = c("metric", "sidecar"),
  entities = list(subject = "01", datatype = "anat")
)
ngeo_read_artifact_batch("derivatives/neurogeo/sub-01")
```

Artifacts use root-relative paths and content SHA-256 values. Batch scope is
fixed to `derivative_only`; neurogeo does not discover or orchestrate an
entire BIDS dataset.
