---
title: "Scalable values, CIFTI, and BIDS derivatives"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scalable values, CIFTI, and BIDS derivatives}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

Delayed storage remains one logical element-by-map values block.

```{r}
backing <- matrix(seq_len(30), nrow = 10, ncol = 3)
delayed <- ngeo_delayed_values(
  function(rows, columns) backing[rows, columns, drop = FALSE],
  dim(backing),
  map_names = c("a", "b", "c"),
  source = "documented callback"
)
x <- ngeo_points(cbind(x = 1:10, y = 0), values = delayed)
chunks <- ngeo_value_chunks(
  x, chunk_size = 4,
  FUN = function(block, rows) colSums(block)
)
chunks
```

A block map is a partition of one logical sparse operator, not a different
mapping. Its hash equals the monolithic map after materialization.

```{r}
atlas <- ngeo_atlas_map(
  x,
  rep(c("A", "B"), each = 5),
  source_support = rep(1, 10)
)
blocked <- ngeo_block_support_map(
  atlas,
  row_block_size = 1,
  source_block_size = 4
)
identical(
  ngeo_support_map_hash(ngeo_materialize_support_map(blocked)),
  blocked$logical_hash
)
```

`write_ngeo_cifti()` writes dscalar, dlabel, and dtseries directly as
CIFTI-2/NIfTI-2 without Workbench. `write_ngeo_bids_derivative()` adds a
scoped JSON sidecar with spatial entities, measurement semantics, and
provenance; it does not index or run a BIDS dataset.
