---
title: "Scalable values, CIFTI, and BIDS derivatives"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scalable values, CIFTI, and BIDS derivatives}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Dense neuroimaging matrices can exceed available memory even when an analysis
needs only a few layers or a block of elements. Chunking changes how values are
accessed, not what the dataset means: element order, map metadata, base
identity, and measurement semantics must remain identical to an in-memory
object.

Chunked access operates on the same logical element-by-map values block.
Standard neuroimaging files use the format-specific file-backed readers.

## Chunked computation

```{r}
backing <- matrix(seq_len(30), nrow = 10, ncol = 3)
colnames(backing) <- c("a", "b", "c")
x <- ngeo_point(
  cbind(x = 1:10, y = 0),
  values = backing,
  measures = do.call(
    rbind,
    replicate(
      3,
      ngeo_measure(support_behavior = "intensive"),
      simplify = FALSE
    )
  )
)
chunks <- ngeo_value_chunks(
  x, chunk_size = 4,
  FUN = function(block, rows) colSums(block)
)
chunks
```

## Bounded support operations

One sparse support map is used directly. Resource budgets bound the values
materialized by the scientific operation; there is no second block-map class.

```{r}
atlas <- ngeo_atlas_map(
  x,
  rep(c("A", "B"), each = 5),
  source_support = rep(1, 10)
)
aggregate_to(
  x,
  atlas$target,
  atlas,
  budget = ngeo_resource_budget(
    memory_bytes = 1024^2,
    materialized_elements = 6
  )
)
```

## Derivative output

`write_ngeo_cifti()` writes dscalar, dlabel, and dtseries directly as
CIFTI-2/NIfTI-2 without Workbench. `write_ngeo_bids_derivative()` adds a
scoped JSON sidecar with spatial entities, measurement semantics, and
history; it does not index or run a BIDS dataset.
