---
title: "Scalable I/O for vertex case-control surfaces"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scalable I/O for vertex case-control surfaces}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
vertex <- ngeo_tutorial_vertex_case_control(n_per_group=5)
```

## One logical brain matrix, bounded access

The teaching matrix has 672 cortical vertices by 10 subjects; the same
contract applies to 32k/164k surfaces and dense CIFTI cohorts. Chunking changes
how values are accessed, not vertex order, subject metadata, base identity, or
measurement semantics.

```{r brain, fig.width=9, fig.height=4.5, fig.alt="Vertex case-control thickness contrast on a bilateral cortical surface", fig.cap="The spatial meaning stays fixed while value access is chunked."}
plot(vertex$difference, chart="flat", show_edges=FALSE, palette="Blue-Red 3")
```

```{r chunks}
chunks <- ngeo_value_chunks(
  vertex$surface, chunk_size=96,
  FUN=function(block, rows) {
    c(first_vertex=min(rows), last_vertex=max(rows),
      HC_sum=sum(block[, vertex$group == "HC", drop=FALSE]),
      SCZ_sum=sum(block[, vertex$group == "SCZ", drop=FALSE]))
  }
)
chunks
```

The callback receives aligned cortical rows. Never reorder a chunk by value or
join subjects by an approximate name: element IDs and layer IDs are the stable
keys.

## Sparse support operations under a budget

```{r support}
atlas <- ngeo_atlas_map(
  vertex$difference, vertex$supports$Schaefer200,
  source_support=ngeo_vertex_area(vertex$difference)
)
regional <- aggregate_to(
  vertex$difference, atlas$target, atlas,
  budget=ngeo_resource_budget(
    memory_bytes=8 * 1024^2,
    materialized_elements=200
  )
)
regional
```

The sparse 200-by-672 operator is consumed directly. The budget cannot silently
downsample the cortex or select another atlas.

## Neuroimaging derivative output

`write_ngeo_cifti()` writes dscalar, dlabel, and dtseries as CIFTI-2/NIfTI-2.
`write_ngeo_bids_derivative()` adds a scoped JSON sidecar containing spatial
entities, measure semantics, and history. For surfaces, use the format-specific
GIFTI/FreeSurfer writers and preserve the vertex-index contract.

```{r derivative, eval=FALSE}
write_ngeo_bids_derivative(
  vertex$difference,
  path="derivatives/neurogeo/group_scz-vs-hc_desc-thickness.dscalar.nii",
  sidecar=list(
    Analysis="SCZ minus HC",
    SamplingUnit="subject",
    Support="registered cortical vertices"
  )
)
```

The writers publish declared derivatives; they do not discover, preprocess, or
run a complete BIDS dataset.
