---
title: "Schema validation and portable manifests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Schema validation and portable manifests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Schema validation checks structural claims before a scientific operation is
allowed to use them. For a spatial dataset this includes domain type,
element/value alignment, measurement metadata, space identity, and the
internal consistency of topology and maps. Validation cannot prove that an
experimental design is scientifically appropriate; it proves that the object
satisfies its declared contract.

NGCS invariants are enforced by one public validation entry point. The schema
table is an internal implementation detail rather than a second runtime
object model.

## Validate an object

```{r}
x <- ngeo_points(
  cbind(x = c(0, 1, 0), y = c(0, 0, 1)),
  values = cbind(signal = c(1, 2, 3)),
  measures = ngeo_measure(spatial_semantics = "intensive"),
  space = ngeo_space("reference-points", kind = "unknown")
)
ngeo_validate(x)
```

## Portable manifests

A portable manifest binds the normative metadata using canonical JSON and
SHA-256. It contains no duplicate values block.

```{r}
manifest <- ngeo_object_manifest(x)
manifest
ngeo_validate_manifest(manifest, x)
```

Atomic JSON exchange verifies integrity before use.

```{r eval=FALSE}
output <- write_ngeo_manifest(manifest, "x.ngcs-manifest.json")
restored <- read_ngeo_manifest(output$path)
```

The public namespace is the authoritative API contract. A manifest is an
integrity and exchange record, not a substitute for the source values or a
workflow engine.
