---
title: "Explicit spaces and transform paths"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explicit spaces and transform paths}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

A coordinate is interpretable only relative to a coordinate reference coordinate_space.
Two files can both use millimetres and still refer to different subjects,
templates, orientations, or surface registrations. A transform is a directed,
auditable claim that coordinates in one exact coordinate_space can be mapped to another.

Matching names are not proof that coordinate spaces are equivalent. A
registry stores the complete normative coordinate_space description and resolves
aliases to one exact hash.

## Register exact spaces

```{r}
native <- ngeo_coordinate_space("native", template = "subject-01")
intermediate <- ngeo_coordinate_space("intermediate", template = "study-template")
standard <- ngeo_coordinate_space("standard", template = "public-template")
registry <- ngeo_coordinate_space_registry(
  list(native, intermediate, standard),
  aliases = c(subject = "native", reference = "standard")
)
ngeo_coordinate_space_audit(native, standard)
```

## Build and resolve a transform graph

The graph contains supplied transforms only. A path composes directed edges;
path resolution does not estimate a missing registration.

```{r}
translation <- function(from, to, shift, method) {
  matrix <- diag(4)
  matrix[1:3, 4] <- shift
  ngeo_transform(
    from,
    to,
    "affine",
    method = method,
    parameters = list(matrix = matrix)
  )
}
native_to_intermediate <- translation(
  native, intermediate, c(1, 0, 0), "supplied step 1"
)
intermediate_to_standard <- translation(
  intermediate, standard, c(0, 2, 0), "supplied step 2"
)
graph <- ngeo_transform_graph(
  registry,
  list(native_to_intermediate, intermediate_to_standard),
  edge_ids = c("native-study", "study-standard")
)
path <- ngeo_transform_path(graph, "subject", "reference")
ngeo_transform_path_history(path)
```

## Apply a path

Path application is geometry-only and requires explicit authorization.
Lossy or unsupported non-affine paths remain auditable but cannot be applied.

```{r}
x <- ngeo_point(
  cbind(x = c(0, 1), y = c(0, 1)),
  values = cbind(signal = c(1, 2)),
  coordinate_space = native
)
changed <- ngeo_apply_transform_path(x, path, authorize = TRUE)
changed$base$geometry$coordinates
```
