---
title: "Bounded spatial modelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bounded spatial modelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  fig.retina = 2,
  out.width = "100%"
)
library(neurogeo)
```

## Background

Spatial models account for dependence that ordinary independent-error
regression ignores. Variograms describe how dissimilarity changes with
distance, kriging predicts from that covariance structure, GWR estimates
location-specific coefficients, and SAR/SEM/CAR models encode dependence
through a spatial_weights matrix. These methods answer different questions and should
not be treated as interchangeable corrections.

Covariance and likelihood calculations are bounded, and each model records
the distance_method, spatial_weights, numerical method, and diagnostics used.

## Data and spatial structure

```{r}
coordinates <- as.matrix(expand.grid(x = 0:4, y = 0:4))
predictor <- coordinates[, 1] + coordinates[, 2]
x <- ngeo_point(
  coordinates,
  values = cbind(
    response = 1 + 2 * predictor + sin(predictor),
    predictor = predictor
  )
)
spatial_weights <- ngeo_spatial_weights(
  x, method = "distance_band", threshold = 1.01, style = "W"
)
```

```{r model-input-map, fig.alt="A five by five point grid colored by response", fig.cap="Always inspect the response on its base before fitting a spatial model."}
plot(
  x,
  layer = "response",
  palette = "Inferno",
  cex = 2.2
)
```

## Variogram and kriging

```{r}
empirical <- ngeo_variogram(x, "response", breaks = 6)
variogram <- ngeo_fit_variogram(empirical, model = "exponential")
variogram
ngeo_kriging(x, "response", variogram, targets = 1:3, neighbors = 8)
```

```{r variogram-plot, fig.alt="Empirical semivariance plotted against spatial distance", fig.cap="The empirical variogram makes the distance scale of spatial dependence visible."}
plot(empirical, pch = 21, bg = "#26a69a", col = "white", cex = 1.5)
```

## Geographically weighted regression

```{r}
bandwidth <- ngeo_gwr_bandwidth(
  x, "response", "predictor",
  candidates = c(1.5, 2.5, 4.5),
  cv = "kfold", folds = 5L
)
gwr <- ngeo_gwr(x, "response", "predictor", bandwidth)
gwr[, c("element_id", "fitted", "condition_number")]
```

`cv = "kfold"` constructs spatially contiguous, element-ID keyed validation
blocks. It does not assign folds from row position. Reordering the same base
therefore cannot change the selected bandwidth or CV loss.

```{r gwr-diagnostics, fig.width=8, fig.height=4.5, fig.alt="Maps of GWR fitted values and condition numbers over the five by five grid", fig.cap="Fitted values must be read together with local numerical diagnostics."}
old_par <- par(mfrow = c(1, 2), mar = c(4, 4, 3, 1))
fit_colors <- grDevices::hcl.colors(32, "Viridis")[
  cut(gwr$fitted, breaks = 32, labels = FALSE, include.lowest = TRUE)
]
plot(coordinates, pch = 21, bg = fit_colors, col = "white", cex = 2.2,
     asp = 1, xlab = "x", ylab = "y", main = "GWR fitted value")
condition_colors <- grDevices::hcl.colors(32, "YlOrRd", rev = TRUE)[
  cut(gwr$condition_number, breaks = 32, labels = FALSE,
      include.lowest = TRUE)
]
plot(coordinates, pch = 21, bg = condition_colors, col = "white", cex = 2.2,
     asp = 1, xlab = "x", ylab = "y", main = "Condition number")
par(old_par)
```

## Spatial regression and CAR

SAR and SEM use an exact dense log determinant only below a declared resource
limit:

```{r}
sar <- ngeo_spatial_regression(
  x, "response", "predictor", spatial_weights, model = "sar"
)
sar
ngeo_car(x, "response", spatial_weights, type = "proper", precision = 1)
```

For large domains, callers must choose a validated bounded approximation in a
future extension rather than receiving an implicit algorithm change.
