---
title: "Bounded spatial modelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bounded spatial modelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

NGCS 2.4 keeps covariance and likelihood calculations bounded and records
the metric, weights, numerical method, and diagnostics used by each model.

```{r}
coordinates <- as.matrix(expand.grid(x = 0:4, y = 0:4))
predictor <- coordinates[, 1] + coordinates[, 2]
x <- ngeo_points(
  coordinates,
  values = cbind(
    response = 1 + 2 * predictor + sin(predictor),
    predictor = predictor
  )
)
weights <- ngeo_weights(
  x, method = "distance_band", threshold = 1.01, style = "W"
)
```

```{r}
empirical <- ngeo_variogram(x, "response", breaks = 6)
variogram <- ngeo_fit_variogram(empirical, model = "exponential")
variogram
ngeo_kriging(x, "response", variogram, targets = 1:3, neighbors = 8)
```

```{r}
bandwidth <- ngeo_gwr_bandwidth(
  x, "response", "predictor",
  candidates = c(1.5, 2.5, 4.5)
)
gwr <- ngeo_gwr(x, "response", "predictor", bandwidth)
gwr[, c("element_id", "fitted", "condition_number")]
```

SAR and SEM use an exact dense log determinant only below a declared resource
limit:

```{r}
sar <- ngeo_spatial_regression(
  x, "response", "predictor", weights, model = "sar"
)
sar
ngeo_car(x, "response", weights, type = "proper", precision = 1)
```

For large domains, callers must choose a validated bounded approximation in a
future extension rather than receiving an implicit algorithm change.
