---
title: "Spatial modelling of a DK68 SCZ-control cortical contrast"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial modelling of a DK68 SCZ-control cortical contrast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
dk <- ngeo_tutorial_dk_case_control()
model_data <- ngeo_parcellation(
  dk$regions,
  values=cbind(
    SCZ_minus_HC=ngeo_values(dk$difference)[, 1L],
    HC_mean=dk$group_mean[, "HC"]
  ),
  centroid=dk$centroids, support_size=dk$support_size,
  adjacency=dk$adjacency,
  measures=rbind(
    ngeo_measure(name="case-control thickness contrast",
                 support_behavior="intensive", unit="mm"),
    ngeo_measure(name="HC mean thickness",
                 support_behavior="intensive", unit="mm")
  ),
  coordinate_space=ngeo_spatial_base(dk$cohort)$coordinate_space
)
```

## Model a cortical quantity, not a point-grid signal

We ask whether the regional SCZ-minus-HC thickness contrast varies with mean HC
thickness while accounting for spatial structure on DK68. This is a model of
68 parcel estimates. It does not replace the subject-level case-control model
that produced those estimates.

```{r input-map, fig.width=9, fig.height=4.6, fig.alt="DK68 SCZ minus HC cortical thickness contrast", fig.cap="Inspect the response on cortex before choosing a spatial covariance or local model."}
ngeo_tutorial_plot_dk(
  model_data, "SCZ_minus_HC", title="Model response: SCZ minus HC thickness"
)
```

## Diagnose spatial scale

```{r variogram}
empirical <- ngeo_variogram(
  model_data, "SCZ_minus_HC", breaks=8,
  distance_method="region_centroid"
)
empirical
```

```{r variogram-plot, fig.width=7, fig.height=4.5, fig.alt="Empirical semivariance of the DK case-control contrast against parcel centroid distance", fig.cap="The empirical variogram displays distance scale; it does not by itself validate a covariance model."}
plot(empirical, pch=21, bg="#2a9d8f", col="white", cex=1.4)
```

Atlas drawing centroids are adequate for this teaching diagnostic only. A real
cortical model should use registered physical centroids or surface geodesics.
Do not fit kriging merely because a variogram can be drawn; the covariance
family and metric must be scientifically defensible.

## Geographically weighted regression

Bandwidth candidates are derived from the declared centroid distances and
selected with spatially blocked, element-ID-keyed cross-validation.

```{r gwr}
distance <- ngeo_distance(
  model_data, seq_len(68), seq_len(68),
  distance_method="region_centroid"
)
candidates <- as.numeric(stats::quantile(
  distance[distance > 0], c(0.15, 0.30, 0.50)
))
bandwidth <- ngeo_gwr_bandwidth(
  model_data, "SCZ_minus_HC", "HC_mean",
  candidates=candidates, distance_method="region_centroid",
  cv="kfold", folds=5
)
gwr <- ngeo_gwr(
  model_data, "SCZ_minus_HC", "HC_mean", bandwidth,
  distance_method="region_centroid"
)
bandwidth
```

Return both fitted values and numerical diagnostics to cortex.

```{r gwr-maps, fig.width=9, fig.height=4.5, fig.alt="DK brain maps of GWR fitted contrast and local condition number", fig.cap="A local fitted map must be read together with its condition-number map."}
fitted_map <- ngeo_parcellation(
  dk$regions, values=cbind(GWR_fitted=gwr$fitted),
  centroid=dk$centroids, support_size=dk$support_size,
  adjacency=dk$adjacency,
  measures=ngeo_measure(name="GWR fitted contrast",
                        support_behavior="intensive", unit="mm"),
  coordinate_space=ngeo_spatial_base(dk$cohort)$coordinate_space
)
condition_map <- ngeo_parcellation(
  dk$regions, values=cbind(condition_number=gwr$condition_number),
  centroid=dk$centroids, support_size=dk$support_size,
  adjacency=dk$adjacency,
  measures=ngeo_measure(name="GWR condition number",
                        support_behavior="intensive", unit="dimensionless"),
  coordinate_space=ngeo_spatial_base(dk$cohort)$coordinate_space
)
ngeo_tutorial_plot_dk(fitted_map, title="GWR fitted SCZ minus HC")
ngeo_tutorial_plot_dk(condition_map, title="GWR local condition number")
```

## Graph-based spatial regression and CAR

```{r graph-models}
weights <- ngeo_spatial_weights(
  model_data, method="region_contiguity", style="W"
)
sar <- ngeo_spatial_regression(
  model_data, "SCZ_minus_HC", "HC_mean", weights, model="sar"
)
car <- ngeo_car(
  model_data, "SCZ_minus_HC", weights, type="proper", precision=1
)
sar
car
```

GWR, SAR, and CAR are different estimands, not interchangeable corrections.
Report the DK support, graph or metric, normalization, bandwidth/CV rule,
covariates, numerical bounds, and that the modeled observations are regional
contrast estimates rather than independent subjects.
