---
title: "Explicit temporal and spatiotemporal analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explicit temporal and spatiotemporal analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Background

Repeated spatial measurements have two distinct axes: spatial elements and
time-indexed maps. Duplicating geometry at every time point wastes memory and
can obscure whether a transformation affected space, time, or both. Temporal
semantics also matter: an instantaneous value, an interval total, and a rate
cannot be contrasted in the same way.

Time is aligned to maps while the object retains one spatial domain and one
element-by-map values block. Point, vertex, voxel, region, or grayordinate
geometry is not expanded through time.

## Attach a time axis

```{r}
library(neurogeo)

coordinates <- matrix(
  c(0, 0, 1, 0, 2, 0),
  ncol = 2,
  byrow = TRUE
)
values <- matrix(
  c(
    1, 2, 4,
    2, 4, 8,
    4, 8, 16,
    8, 16, 32
  ),
  nrow = 3
)
x <- ngeo_points(coordinates, values = values)
axis <- ngeo_time_axis(
  time = c(0, 1, 3, 6),
  unit = "second"
)
x <- ngeo_set_time_axis(
  x, axis, temporal_semantics = "instantaneous"
)
```

## Slice time

Time slicing changes only the aligned map dimension.

```{r}
slice <- ngeo_time_slice(x, index = c(1, 3))
identical(ngeo_domain_hash(slice), ngeo_domain_hash(x))
ngeo_get_time_axis(slice)$time
```

## Space-time weights

Spatial and temporal operators remain independent sparse components. The
combined operator records a Kronecker-sum contract but does not store the
space-by-time matrix.

```{r}
spatial <- ngeo_weights(
  x, method = "distance_band", threshold = 1.1, style = "W"
)
temporal <- ngeo_temporal_weights(axis, style = "W")
space_time <- ngeo_spatiotemporal_weights(
  spatial, temporal, combination = "sum", spatial_scale = 0.5
)

lag <- ngeo_spatiotemporal_lag(x, space_time)
moran <- ngeo_spatiotemporal_moran(
  x, space_time, permutations = 99, seed = 33
)
moran
```

## Change and trend

Variograms require a hard pair budget. Longitudinal helpers derive ordinary
spatial maps with source domain and axis identities in provenance.

```{r}
ngeo_temporal_variogram(x, breaks = 3, max_pairs = 100)
ngeo_longitudinal_change(x, from = 1, to = 4, scale = "rate")
ngeo_temporal_trend(x)
```

## Interval measurements

For interval data, construct explicit interval boundaries and use compatible
semantics.

```{r}
interval_axis <- ngeo_time_axis(
  time = c(0.5, 2),
  unit = "hour",
  interval_start = c(0, 1),
  interval_end = c(1, 3)
)
interval_data <- ngeo_points(
  coordinates,
  values = matrix(1:6, nrow = 3)
)
interval_data <- ngeo_set_time_axis(
  interval_data, interval_axis, temporal_semantics = "rate"
)
ngeo_temporal_contrast(interval_data, operation = "integral")
```
