---
title: "Support-aware inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support-aware inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

An effect can change when the same source data are summarized with different
atlases or spatial scales. Testing each atlas independently confounds this
support sensitivity with different random null draws and inflates the number
of opportunities for a false positive. A common-support test treats the
declared atlases as one inferential family.

One source-base null realization is generated per replicate and applied to
every support map. The resulting joint null distribution supports family-wise
adjustments such as maxT while preserving the dependence among atlas results.

## Define a support family

```{r}
source <- ngeo_surface(
  coordinates = cbind(
    x = c(0, 1, 2, 0, 1, 2),
    y = c(0, 0, 0, 1, 1, 1),
    z = 0
  ),
  faces = matrix(
    c(1, 2, 5, 1, 5, 4, 2, 3, 6, 2, 6, 5),
    ncol = 3,
    byrow = TRUE
  ),
  values = cbind(
    outcome = c(1.1, 2.9, 5.2, 2.8, 7.1, 10.9),
    predictor = c(0, 1, 2, 1, 3, 5)
  ),
  measures = rbind(
    ngeo_measure(support_behavior = "intensive"),
    ngeo_measure(support_behavior = "intensive")
  ),
  coordinate_space = ngeo_coordinate_space("support-inference-example", kind = "surface")
)
first <- ngeo_atlas_map(source, c("A", "A", "B", "B", "C", "C"))
second <- ngeo_atlas_map(
  source,
  c("A", "B", "B", "C", "C", "A"),
  target = first$target
)
layers <- list(first = first, second = second)
targets <- rep(list(first$target), 2)
```

## Choose a null model

An unconstrained permutation is explicitly marked as non-spatial. Use a
Moran spectral null with matching spatial_weights or a surface spin with valid
spherical registration coordinates when spatial preservation is required.

```{r}
family_test <- ngeo_common_support_test(
  source,
  layers,
  targets,
  outcome = "outcome",
  predictor = "predictor",
  adjustment = "maxT",
  nsim = 99,
  seed = 2301
)
family_test
family_test$estimates
```

## Summarize robustness

Atlas effects include intervals, random-effects consensus, heterogeneity,
and leave-one-atlas-out influence.

```{r}
effect <- ngeo_atlas_robust_effect(
  source,
  layers,
  targets,
  outcome = "outcome",
  predictor = "predictor"
)
effect$estimates
effect$meta_analysis
```

## Multiscale inference

Scale order is never inferred. It must be declared by the caller.

```{r}
multiscale <- ngeo_multiscale_inference(
  source,
  layers,
  targets,
  scales = c("coarse", "fine"),
  outcome = "outcome",
  predictor = "predictor",
  nsim = 49,
  seed = 2302
)
multiscale$adjacent_change
```

These results quantify sensitivity over the supplied support family. They do
not establish parcellation- or scale-invariant local effects.
