---
title: "Support-aware inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support-aware inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

Version 2.3 tests a declared family of spatial supports by generating one
source-domain null realization per replicate and applying that same
realization to every support map.

```{r}
source <- ngeo_surface(
  coordinates = cbind(
    x = c(0, 1, 2, 0, 1, 2),
    y = c(0, 0, 0, 1, 1, 1),
    z = 0
  ),
  faces = matrix(
    c(1, 2, 5, 1, 5, 4, 2, 3, 6, 2, 6, 5),
    ncol = 3,
    byrow = TRUE
  ),
  values = cbind(
    outcome = c(1.1, 2.9, 5.2, 2.8, 7.1, 10.9),
    predictor = c(0, 1, 2, 1, 3, 5)
  ),
  measures = rbind(
    ngeo_measure(spatial_semantics = "intensive"),
    ngeo_measure(spatial_semantics = "intensive")
  ),
  space = ngeo_space("support-inference-example", kind = "surface")
)
first <- ngeo_atlas_map(source, c("A", "A", "B", "B", "C", "C"))
second <- ngeo_atlas_map(
  source,
  c("A", "B", "B", "C", "C", "A"),
  target = first$target
)
maps <- list(first = first, second = second)
targets <- rep(list(first$target), 2)
```

An unconstrained permutation is explicitly marked as non-spatial. Use a
Moran spectral null with matching weights or a surface spin with valid
spherical registration coordinates when spatial preservation is required.

```{r}
family_test <- ngeo_common_support_test(
  source,
  maps,
  targets,
  outcome = "outcome",
  predictor = "predictor",
  adjustment = "maxT",
  nsim = 99,
  seed = 2301
)
family_test
family_test$estimates
```

Atlas effects include intervals, random-effects consensus, heterogeneity,
and leave-one-atlas-out influence.

```{r}
effect <- ngeo_atlas_robust_effect(
  source,
  maps,
  targets,
  outcome = "outcome",
  predictor = "predictor"
)
effect$estimates
effect$meta_analysis
```

Scale order is never inferred. It must be declared by the caller.

```{r}
multiscale <- ngeo_multiscale_inference(
  source,
  maps,
  targets,
  scales = c("coarse", "fine"),
  outcome = "outcome",
  predictor = "predictor",
  nsim = 49,
  seed = 2302
)
multiscale$adjacent_change
```

These results quantify sensitivity over the supplied support family. They do
not establish parcellation- or scale-invariant local effects.
