---
title: "Support uncertainty and operator ensembles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support uncertainty and operator ensembles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(neurogeo)
```

## Background

Changing spatial support propagates uncertainty as well as point estimates.
If source errors are correlated, regional variance depends on the full
covariance structure and on the support operator, not only on element-wise
standard errors. Uncertainty in the operator itself is a separate source of
variation.

Uncertainty must be aligned to the same ordered domain as the values it
describes. Independent variance, sparse or full covariance, and low-rank
shared variation are represented without creating a second assay in the
dataset.

## Propagate value uncertainty

```{r}
source <- ngeo_points(
  cbind(x = 1:6, y = 0),
  values = cbind(signal = c(2, 4, 3, 8, 6, 10)),
  measures = ngeo_measure(spatial_semantics = "intensive"),
  space = ngeo_space("uncertainty-example")
)
atlas <- ngeo_atlas_map(
  source,
  c("A", "A", "B", "B", "C", "C"),
  source_support = c(1, 2, 1, 2, 1, 2)
)
covariance <- ngeo_support_covariance(
  source,
  variance = rep(0.5, 6),
  factor = matrix(rep(0.2, 6), ncol = 1)
)
analytic <- ngeo_support_uncertainty(
  source,
  atlas$target,
  atlas,
  covariance,
  output = "covariance"
)
analytic
```

For intensive measurements, analytic propagation uses the derivative of
the support-normalized ratio. The returned diagonal is scalable; full
target covariance has a configurable size guard.

```{r}
monte_carlo <- ngeo_support_uncertainty(
  source,
  atlas$target,
  atlas,
  covariance,
  method = "monte_carlo",
  nsim = 200,
  seed = 2202
)
monte_carlo
```

## Compare operator alternatives

Alternative operators must have identical ordered source and target
domains:

```{r}
probability <- matrix(
  c(
    1, 0, 0,
    0.8, 0.2, 0,
    0.2, 0.8, 0,
    0, 1, 0,
    0, 0.2, 0.8,
    0, 0, 1
  ),
  nrow = 6,
  byrow = TRUE,
  dimnames = list(NULL, c("A", "B", "C"))
)
soft <- ngeo_probabilistic_atlas_map(
  source,
  probability,
  target = atlas$target,
  source_support = c(1, 2, 1, 2, 1, 2)
)
ensemble <- ngeo_segmentation_ensemble(list(hard = atlas, soft = soft))
sensitivity <- ngeo_support_sensitivity(
  source,
  atlas$target,
  ensemble
)
sensitivity
```

## Interpretation

Between-operator variance measures sensitivity to the supplied alternatives.
It is not automatically a calibrated registration or segmentation
confidence interval.

```{r}
ngeo_support_condition(soft)
ngeo_support_diagnostics(
  soft,
  source_structure = rep(c("left", "right"), each = 3)
)
```
