---
title: "Support uncertainty on a vertex SCZ-control cortical map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support uncertainty on a vertex SCZ-control cortical map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>", message=FALSE,
                      warning=FALSE, fig.align="center")
library(neurogeo)
source(system.file("tutorial-code", "brain-case-study.R",
                   package="neurogeo", mustWork=TRUE))
vertex <- ngeo_tutorial_vertex_case_control(n_per_group=5)
```

## The value and its uncertainty share one cortical base

The source is a 672-vertex SCZ-minus-HC thickness map. Value uncertainty and
operator uncertainty are different: the first describes noisy vertex effects;
the second describes uncertain assignment of vertices to parcels.

```{r brain, fig.width=9, fig.height=4.5, fig.alt="Vertex-level SCZ minus HC thickness contrast on bilateral cortex", fig.cap="Uncertainty must stay aligned to the same ordered cortical base as the plotted effect."}
plot(vertex$difference, chart="flat", show_edges=FALSE, palette="Blue-Red 3")
```

## Propagate correlated vertex uncertainty to DK68-sized parcels

```{r analytic}
vertex_area <- ngeo_vertex_area(vertex$difference)
hard <- ngeo_atlas_map(
  vertex$difference, vertex$supports$DK68,
  source_support=vertex_area
)
value_covariance <- ngeo_support_covariance(
  vertex$difference,
  variance=rep(0.0025, length(vertex_area)),
  factor=matrix(rep(0.01, length(vertex_area)), ncol=1)
)
analytic <- ngeo_support_uncertainty(
  vertex$difference, hard$target, hard, value_covariance,
  output="covariance"
)
analytic
```

For intensive thickness, propagation uses the derivative of the
area-normalized ratio. The low-rank factor represents shared uncertainty and
therefore creates target covariance; 672 independent standard errors alone
would miss it.

```{r monte-carlo}
monte_carlo <- ngeo_support_uncertainty(
  vertex$difference, hard$target, hard, value_covariance,
  method="monte_carlo", nsim=200, seed=2202
)
monte_carlo
```

## Compare hard and uncertain parcel boundaries

The soft operator retains 90% membership in the hard parcel and moves 10% to a
neighboring target. This is a transparent boundary-sensitivity scenario, not a
calibrated segmentation probability.

```{r ensemble}
hard_probability <- as.matrix(Matrix::t(hard$operator))
soft_probability <- 0.9 * hard_probability +
  0.1 * hard_probability[, c(2:ncol(hard_probability), 1L)]
target_id <- ngeo_base_elements(hard$target)$region_id
soft <- ngeo_probabilistic_atlas_map(
  vertex$difference, soft_probability, region_id=target_id,
  target=hard$target, source_support=vertex_area
)
ensemble <- ngeo_segmentation_ensemble(list(hard=hard, soft=soft))
sensitivity <- ngeo_support_sensitivity(
  vertex$difference, hard$target, ensemble
)
sensitivity$summary
```

```{r operator-plot, fig.width=7, fig.height=4, fig.alt="Membership-sum diagnostic for the soft cortical support operator", fig.cap="The probabilistic operator remains complete while distributing membership across boundaries."}
plot(soft, type="membership", col="#e76f51", border="white")
```

Between-operator variance measures sensitivity to the supplied hard/soft
alternatives. It is not automatically a registration confidence interval.
Report value covariance assumptions, operator alternatives, target support,
simulation count, seed, and whether intervals are analytic, Monte Carlo, or
between-operator sensitivity ranges.
