---
title: "Transform-aware resampling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transform-aware resampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Background

Resampling evaluates source measurements on a target base. It requires
three separate declarations: a path connecting the coordinate spaces, an
interpolation rule, and policies for coverage, conservation, missing values,
and uncertainty. Registration estimates a spatial relationship; resampling
uses a relationship that has already been supplied. These are not the same
operation.

The transform graph is connected to support mapping without estimating
registration or executing resampling implicitly.

## Register spaces and transforms

Register exact spaces and supplied transforms first.

```{r, eval=FALSE}
native <- ngeo_coordinate_space("native", kind = "volume")
standard <- ngeo_coordinate_space("standard", kind = "volume")

affine <- diag(4)
affine[1, 4] <- 10
known <- ngeo_transform(
  native,
  standard,
  type = "affine",
  method = "supplied scanner-to-template affine",
  parameters = list(matrix = affine)
)

registry <- ngeo_coordinate_space_registry(list(native, standard))
graph <- ngeo_transform_graph(
  registry,
  transforms = list(known),
  edge_ids = "native-to-standard",
  invertible = FALSE,
  lossy = FALSE
)
path <- ngeo_transform_path(graph, "native", "standard")
```

## Create an inert plan

The source and target are ordinary NGCS objects. The target is a base
template; its values are not used.

```{r, eval=FALSE}
plan <- ngeo_resampling_plan(
  source,
  target,
  path,
  method = "linear",
  coverage = "error",
  conservation = "strict",
  missing = "error",
  unknown = "error",
  uncertainty = "none",
  budget = ngeo_resource_budget(
    memory_bytes = 256 * 1024^2,
    materialized_elements = 2e7
  )
)
```

## Authorize execution

Plan creation is inert. Building a map or resampling requires explicit
authorization.

```{r, eval=FALSE}
support_map <- ngeo_build_resampling_map(
  plan, authorize = TRUE
)
diagnostics <- ngeo_resampling_diagnostics(plan, support_map)

result <- ngeo_resample(plan, authorize = TRUE)
result$data
result$diagnostics
```

Surface plans support nearest and barycentric mapping. Volume plans support
nearest, trilinear (`method = "linear"`), and exact axis-aligned overlap.
The path must be supplied, non-lossy, and affine-applicable. An auditable warp
edge remains useful in a transform graph, but this resampling bridge refuses
to execute it.

Coverage and conservation are different decisions. Coverage controls
geometric contributions outside the target; conservation determines whether
non-unit extensive/count allocation fails or is explicitly normalized.
Unknown measurement semantics fail unless the caller selects an
interpretation.

## Uncertainty

Uncertainty must be declared in the plan.

```{r, eval=FALSE}
uncertain_plan <- ngeo_resampling_plan(
  source,
  target,
  path,
  method = "nearest",
  uncertainty = "value"
)
uncertain <- ngeo_resample(
  uncertain_plan,
  value_variance = source_variance,
  authorize = TRUE
)
uncertain$variance
```

The result dataset still has exactly one target spatial base and one
aligned values block. Variance, the sparse map, diagnostics, and joint
path/plan/map hashes are explicit result components.
