---
title: "NIfTI volume: read to auditable result"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NIfTI volume: read to auditable result}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This workflow is for voxel data with an affine. A voxel index is not a physical
coordinate: inspect the affine and coordinate-space unit before choosing a
metric or neighborhood. Zero is retained as data unless an explicit mask says
otherwise.

The code reads a bundled checksum-pinned NIfTI, inspects base/layer/measure
semantics, runs QC and a six-neighbor Moran analysis, plots one layer, performs
a round trip, and inspects history. Moran's I is descriptive for these voxels;
it is not subject-level inference or a causal effect.

```{r setup, include=FALSE}
code <- file.path("..", "inst", "tutorial-code", "workflow-volume.R")
if (!file.exists(code)) code <- system.file("tutorial-code", "workflow-volume.R", package = "neurogeo")
knitr::read_chunk(code)
```

```{r workflow-volume}
```

If qform and sform disagree, neurogeo preserves both and emits a structured
warning. Resolve that conflict from acquisition/BIDS metadata rather than
silencing it. For large files, use the file-backed reader and a finite
`ngeo_resource_budget()`.
