---
title: "DK68 case-control cohort: subjects are layers, parcels are the base"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DK68 case-control cohort: subjects are layers, parcels are the base}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This workflow uses the common tutorial cohort: DK68 cortical thickness for 100
HC and 100 SCZ subjects. Parcels are spatial elements and subjects are
layers--not the reverse. Group, age, sex, site, and subject identifiers belong
in layer metadata; millimetres and intensive support behavior belong in the
measure.

The code validates the full 68-by-200 object, runs QC, derives a descriptive
SCZ-minus-HC map, evaluates that map on declared parcel adjacency, draws the
effect on the brain, and exports redacted history. The spatial Moran test
describes the contrast map; a population diagnosis claim still requires a
subject-level model and defensible subject exchangeability.

```{r setup, include=FALSE}
knitr::read_chunk(system.file(
  "tutorial-code", "workflow-roi-cohort.R", package = "neurogeo"
))
```

```{r workflow-roi-cohort, fig.width=9, fig.height=4.6, message=FALSE, warning=FALSE, fig.alt="Simulated SCZ minus HC cortical thickness on the DK atlas", fig.cap="ROI-by-cohort data stay interpretable because the 68 parcel rows and 200 subject layers remain explicit."}
```

For confirmatory inference, preserve the 200 subjects as independent units,
adjust the prespecified covariates, declare the exchangeability scheme, and
state the estimand, null, family, support, and uncertainty target before using
`ngeo_group_test()` or an external mass-univariate model.
