---
title: "CIFTI grayordinates: components before statistics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CIFTI grayordinates: components before statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

CIFTI rows can mix cortical vertices and subcortical voxels. Inspect brain
models, structures, source indices, and attached surface geometry before
creating topology. Row order is part of the scientific contract.

This workflow attaches matching four-vertex surfaces to the tiny bilateral
CIFTI fixture, runs QC, constructs component-contiguity weights without edges
between structures, computes Moran's I with an explicit zero policy, plots the
ordered values, writes dscalar CIFTI, and verifies the round trip.

```{r setup, include=FALSE}
code <- file.path("..", "inst", "tutorial-code", "workflow-cifti.R")
if (!file.exists(code)) code <- system.file("tutorial-code", "workflow-cifti.R", package = "neurogeo")
knitr::read_chunk(code)
```

```{r workflow-cifti}
```

The line plot is an ordering diagnostic, not a cortical map. Without verified
surface geometry, surface topology, area, geodesic distance, and cortical
plotting must remain unavailable.
