---
title: "Cortical surface: geometry, topology, and metric"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cortical surface: geometry, topology, and metric}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A surface measurement is aligned to ordered vertices; faces define topology,
while a coordinate set defines a metric embedding. They are related but not
interchangeable. Never attach a metric array whose vertex order is uncertain.

This executable workflow reads GIFTI geometry, constructs one explicit
intensive measure, validates it, builds mesh-contiguity weights, computes and
interprets Moran's I, creates a declared viewing chart, plots, writes GIFTI,
and checks the restored history.

```{r setup, include=FALSE}
knitr::read_chunk("inst/tutorial-code/workflow-surface.R")
```

```{r workflow-surface}
```

The tutorial chart is only a viewing projection; it is not a cortical
registration or an anatomical flattening. Use a registered surface pair for
cross-subject mapping and record its source and direction.
