---
title: "ROI by cohort: subjects are layers, parcels are the base"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ROI by cohort: subjects are layers, parcels are the base}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

For an ROI-by-subject table, parcels are spatial elements and subjects are
layers—not the reverse. Group and subject identifiers belong in layer metadata;
measurement units and support behavior belong in the deduplicated measure.

The workflow reads a bundled cohort table, constructs and validates that model,
runs QC, derives a case-minus-control descriptive map without internal
mutation, analyzes its spatial pattern on declared parcel adjacency, plots it,
and exports redacted history. This six-subject fixture demonstrates data shape;
it cannot support a population or clinical claim.

```{r setup, include=FALSE}
knitr::read_chunk("../inst/tutorial-code/workflow-roi-cohort.R")
```

```{r workflow-roi-cohort}
```

For confirmatory group inference, create a subject-level exchangeability
schedule and state the estimand, independent unit, null, family, support, and
uncertainty target before running `ngeo_group_test()`.
